### 2024-01-05 analysis/supplementary-materials/SOM_datadictionaries.qmd
- First (first) draft of the article.

---
title: "Data dictionaries"
author:
  - Jane Doe:
      correspondence: "yes"
      email: janedoe@fosg.org
      orcid: 0000-0003-1689-0557
      institute:
        - fosg
        - fop
  - John Q. Doe:
      institute: fosg
      orcid: 0000-0003-1689-0558
  - Peder Ås:
      institute: fosg
      orcid: 0000-0003-1689-0559
  - Juan Pérez:
      orcid: 0000-0003-1689-0551
      institute:
        - name: Acme Corporation
  - Max Mustermann:
      orcid: 0000-0003-1689-0552
title-block-published: "Last updated"  
format: 
  pdf:
    classoption: landscape
    reference-doc: "../templates/template.docx" # Insert path for the DOCX file
execute:
  echo: false
  warning: false
  message: false
  comment: "#>"
  fig-path: "../figures/"
  fig-dpi: 600
filters:
  - ../templates/scholarly-metadata.lua
  - ../templates/author-info-blocks.lua
  - ../templates/pagebreak.lua
bibliography: ../paper/references.bib
csl: "../templates/journal-of-archaeological-science.csl" # Insert path for the bib-style
---

# 1. Data dictionaries

## 1.1. Group characterization

## 1.2. Single characterization

## 1.3. Petrographic characterization

# 2. References