### 2024-04-26 analysis/paper/paper.qmd
Pre submission commit_1

---
title: "Chert use and procurement patterns during the Upper Palaeolithic in Southwesternmost Iberia"
author:
  - Joana Belmiro:
      orcid: https://orcid.org/0000-0003-4951-2907
      correspondence: "yes"
      email: jfbelmiro@ualg.pt
      institute:
        - Interdisciplinary Center for Archaeology and the Evolution of Human Behaviour, University of Algarve, Faro, Portugal
  - Xavier Terradas:
      orcid: https://orcid.org/0000-0002-8000-5607
      institute: 
        - Institución Milá y Fontanals de Investigación en Humanidades, Consejo Superior de Investigaciones Científicas, Barcelona, Spain
  - Salvador Dominguez-Bella:
      orcid: https://orcid.org/0000-0003-3892-763X
      institute: 
        - UGEA-PHAM, Departamento de Ciencias de la Tierra, Facultad de Ciencias, Universidad de Cádiz. Puerto Real, Cádiz, Spain
  - João Cascalheira:
      orcid: https://orcid.org/0000-0003-0321-8892
      institute:
        - Interdisciplinary Center for Archaeology and the Evolution of Human Behaviour, University of Algarve, Faro, Portugal
title-block-published: "Last updated" 
format: 
  docx:
    reference-doc: "../templates/template.docx" # Insert path for the DOCX file
execute:
  echo: false
  warning: false
  message: false
  comment: "#>"
  fig-path: "../figures/"
  fig-dpi: 600
number-sections: true
filters:
  - ../templates/scholarly-metadata.lua
  - ../templates/author-info-blocks.lua
  - ../templates/pagebreak.lua
bibliography: references.bib
csl: "../templates/springer-socpsych-author-date.csl" # Insert path for the bib-style
abstract: |
  Analyses of raw materials and the distinction between local/regional and long-distance sources have proven invaluable for understanding the extensive movements, interactions, and social networks during the Upper Palaeolithic (UP) in the Iberian Peninsula. However, unlike other parts of Iberia, research on the management and acquisition of raw materials in the south and west of the Peninsula remains relatively underdeveloped. Despite significant knowledge about the technological practices of Palaeolithic hunter-gatherers from southern Portugal, particularly from studies conducted at the Vale Boi site, there's a noticeable lack of focus on raw materials. This paper presents the first comprehensive characterization of chert raw materials from the Gravettian, Proto-Solutrean, and Solutrean sequences at Vale Boi, using both macroscopic and petrographic techniques. Our study reveals that the majority of chert found at Vale Boi originates locally, within a 20 km radius, while a smaller portion comes from non-local sources, potentially indicating long-distance procurement from central Portugal. Additionally, the identification of cherts resembling those from southern Spain suggests a significant west-east mobility pattern throughout the UP period.
keywords: |
  Upper Paleolithic; Iberian Peninsula; Lithic raw materials; Petrography
---

<!-- This is the format for text comments that will be ignored during renderings. Do not put R code in these comments because it will not be ignored. -->

<!-- With the following code you can access and display values from the yml header above. -->

Keywords: `r rmarkdown::metadata$keywords`

Highlights: `r rmarkdown::metadata$highlights`

<!-- The actual document text starts here: -->

```{r packages}
#| label: callpackages

# Call packages
library(readr)
library(dplyr)
library(stringr)
library(tidyr)
library(ggplot2)
library(scales)
library(janitor)
library(ggpubr)
library(knitr)
library(gtsummary)

```

```{r dbsetup}
#| label: db-setup

# Terrace DB preparation
# Read DBs
T_RM_data <- read_csv(here::here('analysis/data/raw_data/MACRO_IND_TERRACE.csv')) # Terrace RM analysis
context <- read_csv(here::here('analysis/data/raw_data/VB_Terrace_context.csv')) # Terrace Field DB - lithic artefact info
xyz <- read_csv(here::here('analysis/data/raw_data/VB_Terrace_xyz.csv')) # Terrace Field DB - just coordinates

# Df cleaning and uniformizing
cols_to_concat <- c("UNIT", "ID") # Vector to concatenate Unit and ID in one single variable (e.g. A6-100) - to use with both Field databases
context$ID <- str_squish(context$ID) # Remove empty spaces from ID field
xyz$ID <- str_squish(xyz$ID) # Remove empty spaces from UNIT field

# Concatenate UNIT and ID variables in one single variable (e.g. A6-100)
context <- context %>%
  dplyr::rename("UNIT" = "Unit") %>% # Renaming first column
  unite(col='ID', cols_to_concat, sep="-", remove=FALSE) %>% # Unite Unit with Id s«using cols_to_concat
  select(ID, Level, Spit, Code) %>% # Selecting important variables
  distinct(ID, .keep_all = TRUE)

xyz <- xyz %>%
  unite(col='ID', cols_to_concat, sep="-", remove=FALSE) %>% # Unite Unit with Id s«using cols_to_concat
  filter(Suffix == 0) %>% # To avoid several entries related to several points of the same artefact
  select(ID, X, Y, Z) %>% # Selecting important variables
  distinct(ID, .keep_all = TRUE)

field_data <- full_join(xyz, context, by = "ID") # Join field tables
T_dataset <- left_join(T_RM_data, field_data, by = "ID") # Join field and RM type tables

T_dataset <- T_dataset %>% 
  filter(!Level %in% c("2", "3")) %>%
  drop_na(Level) %>% 
  mutate(Level = case_when(Level == "7C" ~ "7",
                           Level == 8 ~ "7", # Collapsing samples from Level 7C and Level 8 into 7, since 8 disappears in field_data
                           Level == "4B" ~ "4", # collapsing samples from 4B to 4 due to small number
                           TRUE ~ Level)) %>% 
  mutate(RMTYPE = case_when(RMTYPE == "Chalcedony" ~ "TYPE 1", # Altering first group names to final group names. Subtypes are maintained as before.
                            RMTYPE == "TYPE 15" ~ "TYPE 3",
                            RMTYPE == "TYPE 7" ~ "TYPE 4",
                            RMTYPE == "TYPE 12" ~ "TYPE 5",
                            RMTYPE == "TYPE 11" ~ "TYPE 6",
                            RMTYPE == "TYPE 13" ~ "TYPE 7",
                            RMTYPE == "TYPE 11F" ~ "TYPE 8",
                            RMTYPE == "TYPE 13B" ~ "TYPE 9",
                            RMTYPE == "Oolitic" ~ "TYPE 10",
                            RMTYPE == "GB" ~ "TYPE 11",
                            TRUE ~ RMTYPE))

# Mutate level 5 to add phases: top of 5 (4E/5) and bottom of 5 (5), following Belmiro et al. (2020).
phase_dataset <- T_dataset %>% 
  mutate(Level = case_when(
    Level == 5 & Z > 24.1 ~ "4E/5",
    Level == 5 & Z < 24.1 ~ "5",
    TRUE ~ Level)) %>% 
  mutate(Technocomplex = case_when(
    Level == 4 | Level == "4C" | Level == "4D" ~ "Solutrean",
    Level == "4E" | Level == "4E/5" | Level == 5 ~ "Proto-Solutrean",
    Level == 6 | Level == 7 ~ "Gravettian"))

# Shelter DB preparation
S_RM_data <- read_csv(here::here('analysis/data/raw_data/MACRO_IND_SHELTER.csv')) # Shelter RM analysis
S_field_data <- read_csv(here::here('analysis/data/raw_data/VB_Shelter_xyz_codes.csv')) # Shelter coordinates
S_layerdata <- read_csv(here::here('analysis/data/raw_data/shelter_layers_JC.csv')) # Shelter layers

S_RM_data <- S_RM_data %>%
  mutate_at("ID", str_replace, "VB", "") %>% # Remove "VB" from ID
  mutate(RMTYPE = case_when(
    RMTYPE == "Type 2" ~ "TYPE 2",
    RMTYPE == "Chalcedony" ~ "TYPE 1", # Altering first group names to final group names. Subtypes are maintained as before.
    RMTYPE == "TYPE 15" ~ "TYPE 3",
    RMTYPE == "TYPE 7" ~ "TYPE 4",
    RMTYPE == "TYPE 12" ~ "TYPE 5",
    RMTYPE == "TYPE 11" ~ "TYPE 6",
    RMTYPE == "TYPE 13" ~ "TYPE 7",
    RMTYPE == "TYPE 11F" ~ "TYPE 8",
    RMTYPE == "TYPE 13B" ~ "TYPE 9",
    RMTYPE == "Oolitic" ~ "TYPE 10",
    RMTYPE == "TYPE GB" ~ "TYPE 11",
    .default = RMTYPE))

cols_to_concat_shelter <- c("YEAR", "ID") # Concatenate YEAR and ID variables in one single variable (e.g. 05-500)

S_layerdata <- S_layerdata %>% 
  mutate_at("Date", str_replace, "20", "") %>%
  dplyr::rename("YEAR" = "Date") %>% 
  unite(col='ID', cols_to_concat_shelter, sep="-", remove=FALSE) %>% 
  select(ID, Layer)

S_field_data <- S_field_data %>%
  mutate_at("YEAR", str_replace, "20", "") %>%
  unite(col='ID', cols_to_concat_shelter, sep="-", remove=FALSE) %>% # Unite YEAR with Id s«using cols_to_concat
  select(ID, X, Y, Z, CODE) %>% # Selecting important variables
  distinct(ID, .keep_all = TRUE)

shelter_field_data <- full_join(S_field_data, S_layerdata, by = "ID")
S_dataset <- left_join(S_RM_data, shelter_field_data, by = "ID", relationship = "many-to-many")

```

# Introduction

Knappable raw materials play a crucial role in understanding the mobility and lifeways of past hunter-gatherers, given the ubiquitous presence of lithic artefacts throughout prehistory. More than mere rocks with which stone tools were produced, the management of these resources is intimately connected to a group’s technological, social, and cultural organization, potentially influencing the group’s overall survivability [@oestmoFormalModelingApproach2017; @binfordOrganizationFormationProcesses1979; @bleedOptimalDesignHunting1986; @bousmanHuntergathererAdaptationsEconomic1993; @torrenceTimeBudgetingHuntergatherer1983; @gouldLithicProcurementCentral1985]. Several key topics in the study of hunter-gatherer behaviour and organization have been explored through raw material analysis, such as modalities of procurement and mobility strategies [@ambrose_social_1990; @binfordOrganizationFormationProcesses1979; @binford_righteous_1985; @gouldEmpiricistStrikesBack1985; @gouldLithicProcurementCentral1985; @kuhnUnpackingReductionLithic1991; @mccall_behavioral_2007], occupation types and their duration [@kuhnUpperPaleolithicRaw2004; @surovellBehavioralEcologyLithic2009], the establishment and dimension of social networks [@whallon_social_2006], as well as exchanges between groups or individuals [@gamblePalaeolithicSocietiesEurope1999]. This is often achieved by systematically characterizing geological and archaeological raw materials and establishing correlations between samples from both origins. The ultimate aim is to pinpoint corresponding sources, enabling a precise differentiation between local, regional, and long-distance raw materials, as well as to examine their distribution within a site over different occupations and periods. These approaches and concepts have also been extensively applied to the study of lithic assemblages from prehistoric archaeological sites in the Iberian Peninsula [@aubry_upper_2016; @aubryEconomyLithicRaw2009; @garcia-rojasGreatStepForward2021; @gomezdesolerPanadellaChertMontmaneu2020; @herrero-alonsoManagementLithicRaw2020; @matias_raw_2016; @ortegaMobilitatDesplacamentsDels2003; @pereiraLithicEconomyTerritory2016; @pereira_raw_2021; @pereira_flint_2022; @ramacciottiMovingLandFirst2022; @sanchezdelatorreNewDataChert2023; @soto_procurement_2016], and contributed to understanding how different groups organized their technology, managed time and risk, and explored the available resources during different chronological periods.

Specifically, analyses of raw materials and the distinction between local/regional and long-distance sources have been instrumental in tracing extensive movements, contacts, and social networks throughout the Upper Palaeolithic (UP). Noteworthy examples include studies in northwestern Iberia (Galicia) during the Solutrean period [@hermidaSilexNWPeninsula2016], northern Iberia (Asturias) during the Magdalenian occupations at Las Caldas [@corchonrodriguezMobiliteTerritoiresRelations2016], northeastern Iberia (Catalunya) during the Aurignacian and Gravettian periods at Arbreda cave [@marreirosExploringLithicVariability2016; @ortegaMobilitatDesplacamentsDels2003], northern Portugal during the Gravettian/Solutrean periods in the Côa Valley [@aubry_upper_2016; @aubryRawMaterialProcurement2004], and inland central Iberia (Guadalajara) during the Proto-Solutrean and Solutrean periods at Peña Capón [@sanchezdelatorreNewDataChert2023]. Although such contacts are commonly recognized and identified in later prehistoric periods [@costa_lithic_2022; @noceteRelacionesCentroPeriferia2005; @rodriguezPrehistoricFlintExploitations2011], the recognition of these connections has broadened our comprehension of the often extensive social networks and contacts during the UP in Iberia, illustrating a vast interconnected territory.

However, in contrast to other areas of the Iberia, and despite the potential of raw material studies to shed light on the lifeways and organizational structures of Palaeolithic hunter-gatherers, research on the management and procurement of raw materials during the UP in southern and western Iberia remains in its initial stages. Specifically, in southern Portugal, even though there is considerable knowledge about the technological organization of Palaeolithic hunter-gatherer groups—largely due to research conducted at the Vale Boi archaeological site [see @bicho_ecodynamics_2013; @cascalheiraHunterGathererEcodynamics2013; @marreirosLithicTechnologyGravettian2015]—studies focusing on raw materials are notably infrequent. This situation is attributed to the absence of a detailed and comprehensive characterization of regional lithic resources in the Algarve region (southern Portugal) and the subsequent creation of a reference collection, which is a critical component for conducting raw material studies. Such efforts are particularly crucial for distinguishing between local and non-local raw materials and for understanding variations within the archaeological record [@pop_simulating_2015; @surovellBehavioralEcologyLithic2009]. Recently, we have made significant strides by introducing a reference collection and establishing a framework for raw material studies in the region, through a detailed macroscopic and petrographic database of chert resources in the Algarve region [@belmiroCreatingFramesReference2023]. This initiative paves the way for systematic investigations into raw material procurement strategies during the Upper Palaeolithic, especially concerning the Vale Boi site.

Vale Boi holds a pivotal role in the study of UP adaptations on the Iberian Peninsula, providing invaluable insights into the subsistence and technological strategies of hunter-gatherer populations [@belmiro_gravettian-solutrean_2021; @bicho_ecodynamics_2013; @cascalheira_2017; @cascalheiraTerritorialityOrganizationTechnology2019; @horta_role_2019; @manneIntensiveSubsistencePractices2012; @marreirosExploringLithicVariability2016; @pereira_territory_2016-1]. The rich archaeological record from the site, indicative of prolonged human occupation, is essential for understanding raw material management practices in UP contexts.

Preliminary studies at Vale Boi, employing macroscopic analysis, identified a predominant use of local cherts from Jurassic limestone formations, complemented by a lesser extent of non-local cherts throughout various UP occupations [@bicho_ecodynamics_2013; @pereira_territory_2016-1; @verissimo_jazidas_2004]. According to Pereira et al. [-@pereira_territory_2016-1], the local cherts likely originated within a 16 km radius, reflecting embedded procurement strategies. In contrast, the presence of non-local cherts points to extended sourcing networks, possibly involving travel to other regions of the Algarve, Alentejo, or southern Spain. The presence of chert from the Cretaceous formations of central Portugal has also been suggested [@bicho_ecodynamics_2013].

These results highlight the potential for systematic raw material studies at the site, employing various methods and approaches to analyze the lithic assemblages from multiple UP occupations. Such research aims to comprehend the behaviours of these hunter-gatherer groups over time: how they navigated and utilized their landscape and lithic resources, their movements across a territory bordered by the ocean to the south and west, a large river to the east and mountain ranges to the north, and the nature of the social networks they established. Vale Boi stands as a key archaeological site for this study since it is currently the only site in southern Portugal with a long-term occupation spanning most of the UP and allowing the exploration of raw material use and mobility trends through time [@cascalheira_2017]. Furthermore, previous studies have highlighted the role of Vale Boi in understanding different patterns of territory exploitation and contact with other regions of the Peninsula throughout the UP. During the Solutrean, Vale Boi has been interpreted as an interface between the several sites of Central Portugal and Southern Spain, making it a key site to identify extended social networks that promoted the exchange of information and culture [@cascalheiraValeBoiAlgarve2013; @cascalheira_2017; @cascalheiraSolutrenseEmPortugal2013].

On the contrary, techno-typological studies from the Gravettian occupations of Vale Boi instead show marked differences between regions and suggest a more limited circulation of people and information in southern Iberia [@marreirosLithicTechnologyVariability2013].

To investigate the applicability of these patterns to the management and circulation of lithic raw materials, and to examine the prevailing hypotheses concerning UP mobility and social network exchanges, this paper presents the first comprehensive characterization of chert raw materials derived from the Gravettian, Proto-Solutrean, and Solutrean (c. 32-19 ka cal BP) sequence at Vale Boi.

## Site description

Vale Boi is located on the western coast of the Algarve region in southern Portugal, within a small valley that extends southward to the Atlantic coast, approximately 2 km away (@fig-map b). It is bordered by limestone outcrops that form rock shelters facing west and southwest [@bichoRockshelterStudiesSouthwestern2007; @bichoEdgeCaseVale2012; @bicho_ecodynamics_2013; @cascalheiraHunterGathererEcodynamics2013; @cascalheira_2017; @manneIntensiveSubsistencePractices2012; @mannePryingNewMeaning2011]. The site extends for more than 10,000 sq meters along the slope of the valley, through which three main areas have been excavated between 2000 and 2019: Slope, Shelter, and Terrace.

The Terrace encompasses occupations from the UP to the Early Neolithic, resulting in the identification of eight main lithostratigraphic units. The UP sequence includes several occupations attributed to the Gravettian (levels 8 to 6) between 32 and 27 ka cal BP, Proto-Solutrean (levels 5 to 4E) between 26 and 24 ka cal BP, and Solutrean (level 4D, 4C, 4C, 4 and 3) between 24 and 20 ka cal BP [@belmiro_gravettian-solutrean_2021; @cascalheiraHunterGathererEcodynamics2013; @cascalheira_2017]. The Shelter area shows four main lithostratigraphic units, with Magdalenian, Solutrean and Gravettian occupations. From these, the Solutrean levels show the most intensive occupation, with three, well-preserved, archaeological horizons (layers C to A) and dated from 24 to 22 ka cal BP [@cascalheiraValeBoiAlgarve2013; @cascalheiraTecnologiaLiticaSolutrense2010; @cascalheira_2017].

```{r}
#| label: tbl-c14
#| tbl-cap: Radiocarbon dates for Vale Boi Upper Paleolithic sequence of the Terrace area and Solutrean of the Shelter area, partially adapted from Cascalheira et al. [-@cascalheira_2017] and Belmiro et al. [-@belmiro_gravettian-solutrean_2021], and including new dates. All dates were calibrated or recalibrated using the IntCal20 and Marine20 curves [@reimerIntCal20NorthernHemisphere2020; @heatonMarine20MarineRadiocarbon2020] in OxCal online v.4.4.4 [@ramseyBayesianAnalysisRadiocarbon2009]. For shells a DR value of 123 ± 58 was used, obtained through the CHRONO Marine Reservoir Database at http://calib.org/marine/.

c14_tbl <- read_csv(here::here('analysis/data/raw_data/C14_VB_calBP.csv'))
c14_20 <- read_csv(here::here('analysis/data/raw_data/IntCal20_Marine20_Oxcal.csv'))
dates_cal <- left_join(c14_tbl, c14_20, by = "Lab-ref")
dates_cal <- dates_cal %>% 
  dplyr::rename("Date Cal BP" = "Concatenate") %>% 
  dplyr::select(Area, Level, Culture, `Lab-ref`, Date, SD, `Date Cal BP`, Material)

kable(dates_cal, "simple")

```

```{r}
#| label: fig-map
#| fig-cap: Iberian Peninsula map with locations and geological formations mentioned throughout the paper. a) Detail of the central Portugal region where the Cenomanian formations with chert nodules are located. b) Detail of southwestern Portugal, with chert and chalcedony bearing formations, outcrops, and location of Vale Boi. c) Detail of eastern Algarve with chert bearing formations. The figure was produced using ArcGIS Pro 3.2.

knitr::include_graphics(here::here("analysis/figures/map_layout.png"))
```

## Regional geological background

The Algarve region is a complex territory, marked by several geomorphic sub-regions and geological units that provide different lithological resources used throughout prehistory. Two of the main geological units which characterize the Algarve are: 1) the Baixo Alentejo Flyschoid Group, part of the South Portuguese Zone [@fernandes_new_2012; @oliveira_carta_1984-1; @onezimeNewGeodynamicInterpretation2003], located in the north sector of the region, and characterized by schists, greywackes and quartzites; and 2) the Algarve basin in the south, which is characterized by limestones, dolomites and marls [@terrinha_bacia_2006]. It is within these limestones and dolomitized limestones from the Triassic to the Upper Jurassic where chalcedony and chert nodules and beds can be found [@manuppellaNoticiaExplicativaFolha1987; @manuppellaNoticiaExplicativaFolha2007; @ribeiro_evolucao_2005; @rocha_noticia_1979; @rocha_noticia_1983; @terrinha_bacia_2006]. Following Luedtke [-@luedtke_archaeologists_1992], we use the term chert to refer to sedimentary rocks composed primarily of microcrystalline quartz, which includes other possible varieties such as jasper, agate, or chalcedony. The term chalcedony is used in its petrographic sense and refers to cherts with a fibrous quartz structure. The availability of regional resources containing cherts and the characterization of their lithologies has already been published by our team [@belmiroCreatingFramesReference2023], and we herein present the summarized main findings. The results showed the existence of abundant chert nodules and beds found in the Lower Jurassic (Carixian) formation and Upper Jurassic (Kimmeridgian) formation, mostly outcropping in the cliff and beach areas of westernmost Algarve, at \~20 km southwest from the site of Vale Boi, and with one single inland outcrop located \~8 km east from the site (@fig-map b). These cherts are frequently opaque with yellow, grey, purple, or red colours, occasionally with bands or laminations, and with frequently visible sponge spicules. Although abundant, these chert nodules are frequently small (\~4-10 cm in length), with the largest nodules being around 20 cm, and frequently characterized by high levels of dolomitization or the presence of fractures. Chert sources were also identified in two formations from the eastern area of the Algarve (@fig-map c). The Middle Jurassic Malhão formation chert outcrops were identified at around 80-100 km east of the site and characterized by chert nodules with similar macroscopic features to those found in western Algarve. The Upper Jurassic Jordana formation chert outcrops were identified at around 90 km east of Vale Boi and characterized by chert nodules with significantly different macroscopic features to those of other formations, also with frequently small sizes (\~5 cm), difficult removal from the parent rock and nonexistence of secondary deposits. Chalcedony outcrops approximately 10 km west of Vale Boi (@fig-map b), within the marl-carbonated Triassic formations as lenses or nodules and was formed through hydrothermal processes (Bicho et al., 2013). The chalcedony frequently has fractures and fissures and is characterized by a fibrous quartz structure, without inclusions or fossils.

# Materials and methods

To characterize the siliceous raw materials used during the UP occupations at Vale Boi and to identify their geological sources, we focused on chalcedony and chert archaeological materials from the Terrace and Shelter areas. These two areas were chosen since both have detailed spatial information, several radiocarbon dates through different levels, and previous studies providing lithic technology, faunal, and geoarchaeological data. Chert was chosen for this study as it is one of the three main used raw materials at the archaeological site of Vale Boi alongside quartz and greywacke. Other raw materials such as chalcedony, dolerite, and schists have also been identified [@belmiro_gravettian-solutrean_2021; @cascalheiraTecnologiaLiticaSolutrense2010; @pereira_territory_2016-1]. Using the field database with all individually coordinated lithic artefacts, we compared the incidence of chert and chalcedony in opposition to the remaining lithics of other raw materials. @fig-rm-comparison shows that for the selected levels of the Terrace area (@fig-rm-comparison a), chert represents only \~15-20% of the lithic assemblage, while in the Shelter area (@fig-rm-comparison b), chert represents between \~50-70% of the lithic assemblage from the Solutrean layers A, B, and C. Based on previous lithic studies in the terrace and shelter areas, the remaining 80% of other raw materials are mostly composed of quartz and greywacke [@belmiro_gravettian-solutrean_2021; @cascalheiraTecnologiaLiticaSolutrense2010; @marreirosPrimeirasComunidadesHomem2009]. Although systematic raw material research has not been conducted on these materials, according to the regional geological context, they have been interpreted as local and readily accessible in the vicinity of the site [@pereira_territory_2016-1]. Frequently, the quartz and greywacke assemblages include large amounts of shattered and/or unknapped chunks or slabs. Previous studies from the terrace area have also shown that when removing shatter and chunks, the percentages of chert become much more representative when compared to greywacke and quartz [@belmiro_gravettian-solutrean_2021; @belmiroProtoSolutreanLithicTechnology2020]. Based on this, despite the low amounts of individually plotted chert artefacts in the field database, chert continues to be one of the main raw materials to produce lithic stone tools, with complete knapping sequences and formal toolkits. The importance of chalcedony has also been shown in its use to produce formal tool kits during specific technocomplexes at the site [@belmiro_gravettian-solutrean_2021].

```{r}
#| label: fig-rm-comparison
#| fig-cap: "Comparison of chert, chalcedony, and other raw materials piece plotted in the Terrace (a) and Shelter (b) areas of Vale Boi."

# Terrace rm comparison
## Data preparation
context_comp <- field_data %>%
  mutate(Level = case_when(Level == "4B" ~ "4", # Collapsing samples from level 4B into 4
                           Level == "7C" ~ "7",
                           Level == 8 ~ "7", # Collapsing samples from Level 7C and Level 8 into 7
                           TRUE ~ Level)) %>% 
  mutate(Level = case_when( # Dividing 5 as previous
    Level == 5 & Z > 24.1 ~ "4E/5",
    Level == 5 & Z < 24.1 ~ "5",
    TRUE ~ Level)) %>% 
  filter(!Level %in% c("3B", "5B", "5C")) %>% 
  dplyr::filter(Code %in% c("LITHIC", "POINT"))

dataset_comp <- phase_dataset %>% 
  select(ID, RMTYPE)
T_comp <- full_join(context_comp, dataset_comp, by = "ID")
T_comp["RMTYPE"][is.na(T_comp["RMTYPE"])] <- "Other" # All NA's (other RMs) turned into "Other"

T_comp <- T_comp %>% 
 mutate(Rawmaterial = case_when(
    RMTYPE == "Other" ~ "Other",
    RMTYPE == "TYPE 1" ~ "Chalcedony",
    RMTYPE != "Other" | RMTYPE != "TYPE 1" ~ "Chert")) %>% # Collapsing all cherts into one single category
  dplyr::filter(Level %in% c("4", "4C", "4D", "4E", "4E/5", "6", "7")) # Choosing only UP levels (excluding 3)

## Barplot
T_comp_bar <- T_comp %>%
  dplyr::group_by(Level, Rawmaterial) %>% 
  dplyr::summarise(n = n()) %>%
  mutate(percent = n/sum(n)) %>% 
  ggplot(aes(x=Level, y = percent, fill = Rawmaterial)) +
  geom_bar(stat = "identity", width = 0.5) +
  scale_y_continuous(labels = scales::percent_format()) +
  theme_minimal() + #type of theme
  labs(x = "Terrace levels", y = "Count %") +# change labels +# change labels
  guides(fill=guide_legend(title=NULL)) + # remove legend title
  geom_text(aes(label=ifelse(percent >= 0.07, paste0(sprintf("%.1f", percent*100),"%"),"")),
                position=position_stack(vjust=0.5), colour="white", size=3) +
  theme(axis.title.x = element_text(size = 10, vjust = -1))+
  scale_fill_manual(values = c("#C74B50", "#D49B54", "#845460", "#EAD3CB", "#BDC7C9",
                              "#2B4F60", "#BDD2B6", "#798777", "#FCF0C8"))

# Shelter rm comparison
## Data preparation
S_field_comp <- S_field_data %>% 
  filter(CODE == "L") # Filtering only lithics
S_data_comp <- S_dataset %>% 
  dplyr::select(ID, RMTYPE)
S_comp <- full_join(S_field_comp, S_data_comp, by = "ID")
S_layerdata <- S_layerdata %>% 
  distinct() # Keeping only individual IDs to avoid repetitions
S_comp <- left_join(S_comp, S_layerdata, by = "ID")
S_comp["RMTYPE"][is.na(S_comp["RMTYPE"])] <- "Other" # All NA's (other RMs) turned into "Other"
S_comp["Layer"][is.na(S_comp["Layer"])] <- "No info" # Unknown Shelter levels turned into "No info" to avoid NA

S_comp <- S_comp %>% 
 mutate(Rawmaterial = case_when(
    RMTYPE == "Other" ~ "Other",
    RMTYPE == "TYPE 1" ~ "Chalcedony",
    RMTYPE != "Other" | RMTYPE != "TYPE 1" ~ "Chert")) %>% # Collapsing all cherts into one single category
  dplyr::mutate(Layer = case_when(
    Layer == "c" ~ "C",
    TRUE ~ Layer)) %>% 
  filter(!Layer %in% c("D", "Z")) # Choosing only Solutrean occupation layers (A, B and C)

## Barplot
S_comp_bar <- S_comp %>%
  dplyr::group_by(Layer, Rawmaterial) %>% 
  dplyr::summarise(n = n()) %>%
  mutate(percent = n/sum(n)) %>% 
  ggplot(aes(x=Layer, y = percent, fill = Rawmaterial)) +
  geom_bar(stat = "identity", width = 0.5) +
  scale_y_continuous(labels = scales::percent_format()) +
  theme_minimal() + #type of theme
  labs(x = "Shelter levels", y = "Count %") +# change labels
  guides(fill=guide_legend(title=NULL)) + # remove legend title
  geom_text(aes(label=ifelse(percent >= 0.07, paste0(sprintf("%.1f", percent*100),"%"),"")),
                position=position_stack(vjust=0.5), colour="white", size=3) +
  theme(axis.title.x = element_text(size = 10, vjust = -1)) +
  scale_fill_manual(values = c("#C74B50", "#D49B54", "#845460", "#EAD3CB", "#BDC7C9",
                              "#2B4F60", "#BDD2B6", "#798777", "#FCF0C8"))

## Unused barplot: here to understand how the chert ratio changes when A, B and C are collapsed since "No info" has most of the other rms.
#S_comp %>% 
#  dplyr::group_by(Rawmaterial) %>% 
#  dplyr::summarise(n = n()) %>%
#  mutate(percent = n/sum(n)) %>% 
#  ggplot(aes(x=Rawmaterial, y = percent, fill = Rawmaterial)) +
#  geom_bar(stat = "identity", width = 0.5) +
#  scale_y_continuous(labels = scales::percent_format()) +
#  theme_minimal() + #type of theme
#  labs(x = "Shelter levels", y = "Count %") +# change labels
#  guides(fill=guide_legend(title=NULL)) + # remove legend title
#  geom_text(aes(label=ifelse(percent >= 0.07, paste0(sprintf("%.1f", percent*100),"%"),"")),
#                position=position_stack(vjust=0.5), colour="white", size=3) +
#  theme(axis.title.x = element_text(size = 10, vjust = -1)) +
#  scale_fill_manual(values = c("#C74B50", "#D49B54", "#845460", "#EAD3CB", "#BDC7C9",
#                              "#2B4F60", "#BDD2B6", "#798777", "#FCF0C8"))

## Terrace and shelter barplots
ggarrange(T_comp_bar, S_comp_bar, labels = c("a", "b"), nrow = 2, common.legend = TRUE, legend = "bottom")

```

We employed both macroscopic and petrographic methods to achieve a more comprehensive characterization of the materials, aiming to offset the inherent limitations of each method [@luedtke_archaeologists_1992].

Macroscopic analysis, while revealing several limitations related to the absence of quantitative variables [@bustillo_is_2009], proved to be advantageous given the extensive sample size of chalcedony and chert lithics under examination (n=5318). The distinct features of various local cherts facilitate their differentiation from cherts of neighbouring origins. Consequently, a macroscopic approach offered a non-destructive, efficient, and cost-effective method for characterizing all materials [@bustillo_is_2009; @tarrino_materias_2013]. Therefore, we implemented an initial phase of macroscopic analysis, subsequently augmented by petrographic study to diminish the subjectivity of the analysis and enhance the material characterization with petrographic data. The synergistic application of these two methods has been previously validated in the characterization of Algarve region cherts, following recent prospection efforts [@belmiroCreatingFramesReference2023]. Hence, an extensive macroscopic and petrographic collection is available for comparative analysis with archaeological materials, employing consistent methodologies.

This reference collection is the Lusolit lithotheque, hosted physically at ICArEHB (University of Algarve, Faro) and digitally at www.lusolit.icarehb.com, and aided in the characterization and source attribution of cherts. Hand samples were used throughout the macroscopic characterization for comparison with the archaeological materials, and the archaeological thin sections were compared with the previously studied geological chert thin sections. To identify the source of other lithologies which are not congruent with the regional cherts and identify possible long-distance chert procurement, we used hand samples and thin sections of cherts from central Portugal also available in the Lusolit lithotheque. The lithotheque from the Unit of Geoarchaeology and Archaeometry Applied to Historic Artistic and Monumental Heritage from the University of Cadiz (UCA) was also visited to understand the macroscopic variability of the local and regional cherts of the lower Guadalquivir basin, within the Atlantic strip of south-western Spain (present-day provinces of Huelva and Cádiz). Following previous studies, we classify the distance of chert sources from the site as local (1-30 km from the site), regional (30-120 km), and long-distance (\>120 km) [@herrero-alonsoManagementLithicRaw2020; @tarrinoFlintRawMaterial2015].

The macroscopic analysis of the archaeological materials was divided into two steps. The first step included a preliminary characterization using a hand lens of 10x magnification. This step allowed us to create types and subtypes based on similar macroscopic characteristics which have been used in previous studies to describe and discern between different types of chert; these include characteristics such as colour, translucency, and feel [@luedtke_archaeologists_1992], and a data dictionary for the recorded variables can be found in the Supplementary Information (Online Resource 1). In the second step, we used a Nikon SMZ25 stereomicroscope to observe each sample in further detail, which allowed us to better characterize the artefacts (especially regarding inclusions and alterations), and the previously established types. The data dictionary used for the individual samples can also be found in the Supplementary Information (Online Resource 1). Given the inherent macroscopic variability of the geological samples and archaeological specimens, as well as the number of samples, the database used for the individual analysis is simplified and focused on collecting data related to weight, cortex, and alterations. This first step was applied to a sample of 3627 chert artefacts from the UP sequence of the Terrace area (levels 8 to 4) and 2805 chert artefacts from the Solutrean sequence of the Shelter area (layers C to A).

While the advantages of employing a macroscopic methodology for analysing the cherts of Vale Boi are notable, a limitation of this approach is its inability to unequivocally characterize and categorize artefacts that exhibit surface alterations typical of an open-air site, including those related to fire, since weathered and altered surface may destroy visible structures and even alter the chemical composition [@delluniversita_development_2019]. Other studies frequently exclude samples that display intense fire damage or several surface alterations [@delluniversita_development_2019; @gomezdesolerPanadellaChertMontmaneu2020; @soto_procurement_2016]. For the present study, whenever artefacts displayed extensive alterations and could not be unequivocally characterized (@fig-terrace-alterations), they were analysed using the individual sample dataset, but grouped in a category of indeterminate cherts (INDET).

As seen in @tbl-indet-freq, level 4 shows the higher percentages of indeterminate artefacts, corresponding to 40% of the total sample (n=1027). This is due to the frequent alterations of the sample, especially related to patinas and alteration rinds, pits, and fire/heat alterations. For all other levels in the Terrace, indeterminate represents on average \~28% of the chert artefacts, while in the Shelter the frequency of INDET is slightly lower (\~20-25%).

```{r}
#| label: fig-terrace-alterations
#| fig-cap: Samples with extensive alterations which prevented the identification of chert type. a) artefact with cracks, white patina and reddening covering 100% of the surface; b) artefact with surface alterations, displaying a portion with patina (brown) and a portion without patina (grey), cleaned with Hydrochloric acid.

knitr::include_graphics(here::here("analysis/figures/terrace_alterations.png"))

```

```{r}
#| label: tbl-indet-freq
#| tbl-cap: N and percentage of grouped and indeterminate (INDET) artefacts by level in the Terrace (levels 4 to 7) and Shelter (A to C) areas.

terrace_indet_table <- phase_dataset %>% 
  mutate(Group = if_else(RMTYPE=="INDET", "Indeterminate", "Grouped")) %>%
  tabyl(Group,Level) %>% 
  adorn_totals("row") %>%
  adorn_percentages("col") %>%
  adorn_pct_formatting(2) %>%
  adorn_ns("front")

shelter_indet_table <- S_dataset %>% 
  mutate(Group = if_else(RMTYPE=="INDET", "Indeterminate", "Grouped")) %>% 
  mutate_at(vars(Layer), ~replace_na(., "ND")) %>% 
  filter(!Layer %in% c("D", "ND")) %>% 
  tabyl(Group,Layer) %>% 
  adorn_totals("row") %>%
  adorn_percentages("col") %>%
  adorn_pct_formatting(2) %>%
  adorn_ns("front")

indet_table <- full_join(terrace_indet_table, shelter_indet_table, by = "Group")

kable(head(indet_table))

```

To better characterize the different chert types identified in the macroscopic analysis and aid in minimizing the inherent caveats of macroscopic methods, twenty-two thin sections were prepared at the Thin Section Lab (Toul, France). The samples chosen for this analysis include at least one sample from each type present throughout the stratigraphy of the Terrace area, from different archaeological levels (@tbl-TS-list). For types that showed high macroscopic variability (e.g. translucency and colour variation or differentiated bioclast facies), thin sections from sub-types were analysed to collect petrographic data that reflected this variability. Petrographic thin sections were analysed by polarized light microscopy (Nikon LV100ND) to determine the mineral composition, textural characteristics, allochems, and bioclasts, as well as any alterations.

```{r}
#| label: tbl-TS-list
#| tbl-cap: List of archaeological thin sections, with the summarised description of observed macroscopic variability and sub-types.

TS_list <- read_csv(here::here('analysis/data/raw_data/TS_short_list.csv'))

kable(TS_list)

```

The complete petrographic analysis for all thin sections studied in this paper (including macroscopic and microscopic figures) can be found as Supplementary Information Materials at our online repository (++). A detailed macroscopic and petrographic description and figures for each type and sub-type of chert and chalcedony (Online Resource 2) and the data dictionary used for the petrographic analysis (Online Resource 1) can be found in Supplementary Information. The complete set of images of the geological thin sections can also be consulted in the LusoLit online database.

The entirety of the R code used for the analysis, datasets, and visual representations contained in this paper can be accessed through our online research compendium (https://doi.org/10.17605/OSF.IO/DBFT2). We used the rrtools package by Marwick et al. [-@marwickPackagingDataAnalytical2018] to create a research compendium and write a reproducible journal article. The provided files include the complete set of raw data used in the analysis, along with a custom R project (Wickham, 2015) containing the code required to generate all tables and figures. To enable maximum reuse, the code is made available under the MIT license, data under CC‐0, and figures under CC‐BY [additional details can be found in @marwickComputationalReproducibilityArchaeological2017].

# Results

```{r}
#| label: t-table

T_table <- phase_dataset %>%
  filter(RMTYPE !="INDET") %>% 
  mutate(RMTYPE = factor(RMTYPE, levels = c("TYPE 1","TYPE 2", "TYPE 3", "TYPE 4",
                                        "TYPE 5", "TYPE 6","TYPE 7", 
                                        "TYPE 8", "TYPE 9", "TYPE 10", "TYPE 11", "TL"))) %>%
  tabyl(RMTYPE, Level) %>%
  adorn_totals(c("row", "col")) %>%
  adorn_percentages("col") %>%
  adorn_pct_formatting(2) %>%
  adorn_ns("front")

```

## Characterization and source attribution

Through the macroscopic and petrographic analysis, we identified 10 main chert/chalcedony types, present throughout the stratigraphy in the Terrace and Shelter areas (a summarised characterization table can be found in the Supplementary Information, Online Resource 3). A distinct type was identified exclusively in the Shelter area, with only a limited presence in the Terrace (n=2), and 16 varieties, each consisting of one to three samples, were identified across various levels of both the Terrace and Shelter. The latter were clustered into a category named Trace Lithotypes (TL). Through the comparison with the reference collections, we identified cherts congruent with the local sources and cherts that are likely not local due to significant differences in appearance, fossil content, and petrography when compared to the cherts available in the geological setting located within a 30 km radius of Vale Boi.

### Cherts congruent with local sources

The raw materials which are congruent with local sources are the chalcedony and massive, micro-cryptocrystalline cherts (@fig-local-types). The chalcedony (Type 1; @fig-local-types a) is identified as a massive, fibrous variant devoid of allochems (@fig-local-types a2). It represents 6.6% of the overall Terrace cherts (excluding INDET) and 2.6% in the Shelter. It presents a colourless and translucent appearance, frequently exhibiting partial or complete white patination (@fig-local-types a1). This type is marked by extensive crystallizations, irregularities, and fractures, coupled with post-depositional alterations that challenge the identification and preservation of the cortex. Both macroscopic and petrographic analyses confirm its congruence with hydrothermal chalcedony located within lenses in Triassic volcanic formations, which are similarly marked by irregularities and fractures.

```{r}
#| label: tbl-cortex_tab
#| tbl-cap: Frequency of cortex type and thickness by raw material from the Terrace area.

# Cortex table
## Preparing cortex db Terrace
cortex_tab <- phase_dataset %>% 
  select(RMTYPE, CORTEX, CORTEXAPPEARANCE, CORTEXTHICKNESS) %>%
  filter(CORTEX == "YES") %>% # filtering only samples with cortex
  filter(RMTYPE != "INDET") %>% 
  filter(RMTYPE != "TYPE 11") %>% # removing Type 11 since it barely has samples
  mutate(RMTYPE = factor(RMTYPE, levels = c("TYPE 1","TYPE 2", "TYPE 3", "TYPE 4",
                                        "TYPE 5", "TYPE 6","TYPE 7", 
                                        "TYPE 8", "TYPE 9", "TYPE 10", "TL")))

cortex_tab <- cortex_tab %>% # dropping the caps and organizing value order
  mutate(CORTEXAPPEARANCE = case_when(
    CORTEXAPPEARANCE == "ROUNDED" ~ "Rounded",
    CORTEXAPPEARANCE == "PEBBLE" ~ "Pebble",
    CORTEXAPPEARANCE == "OUTCROP" | CORTEXAPPEARANCE == "Outcrop" ~ "Irregular",
    CORTEXAPPEARANCE == "OTHER" | CORTEXAPPEARANCE == "Other" | CORTEXAPPEARANCE == "INDETERMINATE" ~ "Indeterminate",
    TRUE ~ CORTEXAPPEARANCE)) %>%
  dplyr::mutate(CORTEXTHICKNESS = case_when(
    CORTEXTHICKNESS == "MEDIUM" ~ "Medium",
    CORTEXTHICKNESS == "THICK" ~ "Thick",
    CORTEXTHICKNESS == "THIN" ~ "Thin",
    TRUE ~ CORTEXTHICKNESS)) %>% 
  mutate(CORTEXAPPEARANCE = factor(CORTEXAPPEARANCE, levels = c("Irregular", "Rounded", "Pebble", "Indeterminate"))) %>%
  mutate(CORTEXTHICKNESS = factor(CORTEXTHICKNESS, levels = c("Thin", "Medium", "Thick"))) %>% 
  drop_na() # removing missing values that may be cortex=NO instead passed as YES

## gt summary table
cortex_tab %>%
  select(RMTYPE, CORTEXAPPEARANCE, CORTEXTHICKNESS) %>% 
  tbl_summary(by=RMTYPE, label = list(CORTEXAPPEARANCE ~ "Cortex type", CORTEXTHICKNESS ~ "Thickness")) %>% 
  modify_header(label ~ "**Variable**") %>% 
  bold_labels()

```

The cherts are massive micro-cryptocrystalline quartz from a marine environment (@fig-local-types b-h). They represent \~60% of the overall Terrace cherts and \~80% of the chert in the Shelter. Macroscopically, these types show variability which resulted in their separation in four different macroscopic types (Type 2 to 5). In general, these are opaque to sub-translucent cherts, with colours varying from brown, yellow, pale red, and grey, with the frequent presence of spicules and iron oxides. The differences are in their colours, nodule morphology, frequency, and variety of bioclasts (Type 2 including triaxon sponge spicules and echinoderm spines), something already identified in the geological samples [@belmiroCreatingFramesReference2023]. Whenever present, the cortex seems to show a mixture of irregular cortex and rounded cortex, both from an outcrop source (@tbl-cortex_tab). Petrographically, however, these cherts are homogeneous, except for a small number of artefacts with a banded variety within type 5. The macroscopic separation was maintained for its possible usefulness in identifying preferred types of nodules or macroscopic varieties through time and for future correlation with technological data. These cherts are characterized by wackestone textures, with accessory amounts of macrocrystalline quartz and fibrous chalcedony, frequently found replacing fossils and occasionally dolomite. Type 5 shows higher percentages of micrite/sparite. In all samples, fossils range from common to very frequent, albeit frequently unidentifiable. When identifiable in thin section, they are sponge spicules (@fig-local-types b2, c2, h2), with the rare occurrence of possible foraminifera (Type 2 and Type 5) and ostracod (@fig-local-types h3) in a sample from type 5. Porosity varies between 10% and 1%, of vug type. Based on the previously described macroscopic and petrographic characteristics of the Lower Jurassic cherts of the Algarve [@belmiroCreatingFramesReference2023], we interpret Types 2, 3, 4, and 5 to be local, as they are congruent with the described characteristics representative of cherts from the Lower Jurassic (Carixian) and Middle Jurassic Malhão formations of the Algarve basin. Types 2 and 3 are congruent with the frequently available brown/yellow and red nodules and beds (@fig-local-types b-d), ranging from uncommon to very frequent fossils. The macroscopic similarities of Types 4 and 5 are more limited (@fig-local-types e-h), as they show similarities to specific nodules, rarely found in the known outcrops, and mostly located in the inland Lower Jurassic outcrops (@fig-map b). As such, we interpret these cherts as local, but from nodules and outcrops that are no longer frequently available or visible in the landscape. This is especially relevant since the inland outcrops are mostly destroyed and can be found in poor preservation conditions, and with small, broken nodules in the surface, frequently altered [@belmiroCreatingFramesReference2023].

```{r}
#| label: fig-local-types
#| fig-cap: Variability of local chert types and main petrographic characteristics. a) Type 1 macroscopic appearance (a1) and thin-section (a2). b) Type 3 macroscopic appearance (b1) and thin-section (b2). c-d) Type 2 macroscopic appearance (c1-d1) and thin-section (c2-d2). e-f) Type 4 macroscopic appearance (e-f1) and thin-section (f2-f3). g-h) Type 5 macroscopic appearance (g-h1) and thin section (h2-h3).

knitr::include_graphics(here::here("analysis/figures/local_grid.png"))

```

### Cherts not congruent with local sources

We identified six chert types with macroscopic and petrographic characteristics that are not congruent with the chert available in known outcrops in western Algarve (@fig-distance-types). Despite their scarce representation in the archaeological record, these raw materials of presumably allochthonous origin are a valuable source of information. Often these lithological types can be considered as clear lithological markers of mobility, which can contribute to support hypotheses about mobility and contacts with neighbouring populations and territories.

```{r}
#| label: fig-distance-types
#| fig-cap: Variability of non-chert types and main petrographic characteristics. a-b) Type 6 macroscopic appearance (a1-b1) and thin-section (a2-b2); b1-b2 correspond to a sample covered 100% in white patina. c-d) Type 7 macroscopic appearance (c1-d1) and thin-section (c2-d2) of peloidal facies (c1-c2) and oolitic facies (d1-d2). e-f) Type 8 macroscopic appearance (e-f1) and thin-section (f2-f3); white and black arrows correspond to foraminifera. g) Type 9 macroscopic appearance (g1) and thin-section (g2). Type 10 macroscopic appearance (h1) and thin-section (h2).

knitr::include_graphics(here::here("analysis/figures/nonlocal_grid.png"))

```

Type 6 (@fig-distance-types a-b) corresponds to a massive cryptocrystalline quartz with a mudstone texture, from a marine environment. It represents \~18% of the overall Terrace cherts and 12% of the chert in the Shelter. This chert is translucent and shows a variety of colours, ranging from brown, red, and grey, which seem to vary in function of the artefact’s thickness. In unaltered samples, visible bioclasts are rare. The cortex is mostly rounded, and rarely pebble-like (FIG), ranging from thin to medium thickness. Two sub-types were individualized based on macroscopic characteristics such as colour, translucency variation, and patterns, although they may simply reflect the use of specific nodules from within Type 6’s variability. Five thin sections were produced to better characterize the macroscopic variability of this type and sub-types. The petrographic analysis shows the samples are characterized by a massive microstructure, mainly composed of cryptocrystalline quartz (between 95-97% in the samples), with accessory fibrous chalcedony found replacing fossils or filling fractures (@fig-distance-types a2-b2), and in some samples macrocrystalline quartz, microcrystalline quartz, and micrite/sparite. Allochems are opaques, iron oxides, and bioclasts. Bioclasts are unidentifiable and poorly preserved, with rare identifiable fossils such as sponge spicules, replaced by fibrous chalcedony and rarely by microcrystalline quartz. Porosity is present between \<1-1%, of vug-type when identifiable.

The similarities between chert artefacts at the UP occupations of Vale Boi and the chert from the Rio Maior region (Estremadura Portuguesa; @fig-map a) have been previously suggested although without systematic studies to ascertain the attribution [@bicho_ecodynamics_2013; @pereira_territory_2016-1]. These Cretaceous cherts, mostly found in secondary deposition settings, are described as frequently translucent, with geodes, colours ranging from yellow to red or grey, and mineralogically homogeneous between them. Petrographic studies from the Estremadura area highlight the presence of iron oxide accumulations and rare fossil ghosts or frequently difficult to see in thin section [@matias_raw_2016]. When identified, sponge spicules and possible rare foraminifera have been identified [@matiasAprovisionamentoMateriasprimasLiticas2012]. Studies of Upper Cretaceous cherts located in the Lisbon area describe cherts with similar macroscopic characteristics (@fig-map a), with micro-cryptocrystalline quartz mudstones with bioclasts such as sponge spicules or bivalve fragments, or wackestone/packstone textures with abundant bioclasts, including ostracods and different types of foraminifers [@jordaoFlintSourcesMobility2022]. These cherts have also been reported as found in highly variable cobble morphologies and macroscopic traits, which is also observable through the reference samples from central Portugal hosted at LusoLit. Alike the archaeological thin sections from Type 6 and congruent with other studies [@jordaoFlintSourcesMobility2022; @matias_raw_2016], the thin sections produced from the Cretaceous cherts hosted at Lusolit show a massive, cryptocrystalline mudstone structure, with uncommon fossils (ghosts), commonly barely visible, iron oxides and low porosity (\<1%). From these cherts, only the dark grey nodules (RT231; SOM) showed identifiable fossils which were rare ostracods. Based on the macroscopic and petrographic similarities to the geological references from Central Portugal and congruence with previous studies from this area, we interpreted Type 6 (and sub-types) as belonging to the macroscopically variable Cretaceous cherts (Upper Cenomanian formations) from central Portugal, between \~215-250 km north of Vale Boi (@fig-map a).

Type 7 (@fig-distance-types c-d) is a massive micro-cryptocrystalline, peloidal/oolitic packstone, from a marine environment and possibly a high-energy, shallow depositional environment, due to the type, sorting, and preservation conditions of the ooids. In the terrace, it represents 10% of the overall Terrace cherts and less than 1% of the chert in the Shelter. It is characterized by white and grey colours. The cortex is from an outcrop, rounded and in lesser frequencies (\<30%) irregular, and varying between medium thickness to thick. This type shows several facies, with different types and concentrations of allochems and bioclasts, such as peloids (@fig-distance-types c1), ooids (@fig-distance-types d1), and sponge spicules. A detailed scheme of the facies can be found in the Supplementary Information (Online Resource 2). The petrography results show this chert has a massive microstructure and packstone texture (@fig-distance-types c2-d2). It is composed mainly of micro-cryptocrystalline quartz (80%), fibrous chalcedony (10%), and micrite (10%). Allochems are opaques, iron oxides, ooids, and peloids. The ooids are poorly preserved and only uncommonly show concentric lamellae in plane-polarized light. Unlike macroscopical observations, bioclasts under the thin section are all unidentifiable, although rare fossils may be bivalve shells. Porosity is variable (\<1-10%). The consultation of the lithotheque hosted at UCA allowed us to identify a subset of cherts with similarities to those from Type 7 from the Middle Subbetic region (Betic Systems, south Spain, @fig-map). These cherts are massive, sub-translucent, peloidal/oolitic packstones. Allochems include very frequent peloids and ooids. The peloids are densely arranged in the samples, more or less visible depending on the alterations to the surface or the thickness. The ooids are poorly sorted, with oval or round shapes, and replaced by quartz. Uncommonly the ooids show a preserved nucleus. The presence of bioclasts varies between samples, ranging between common to rare, and when present/identifiable include sponge spicules (common), foraminifera (rare to uncommon), and echinoderm spines (rare). Similar cherts have been identified and described in previous works, describing grey cherts with peloidal and oolitic facies and occasional foraminifera [@rodriguezPrehistoricFlintExploitations2011]. Despite the nonexistence of thin sections for comparison, the macroscopic similarities between the cherts allow us to interpret Type 7 as coming from the Upper Jurassic formations of the Betic Systems, located in the south and southeastern Iberian Peninsula, at least 250 km east from the site (@fig-map).

Type 8 (@fig-distance-types, e-f) corresponds to a massive micro-cryptocrystalline quartz from a marine environment and shelf/platform depositional environment. It represents \~2% of the overall Terrace cherts and \~5% of the chert in the Shelter. Identified and well-preserved fossils are foraminifera from the Pfenderenidae family (@fig-distance-types e, f2), which are larger benthic foraminifera from marine environments [@boudagher-fadelChapterMesozoicLarger2008]. The association of these foraminifera indicates this chert formed from sediments deposited possibly between the Jurassic to the Cretaceous. Macroscopically, this type is characterized by grey and white colours (@fig-distance-types e-f1). The translucency ranges from sub-translucent to opaque. Whenever present, the cortex is rounded and thin. Petrographically, this type is a packstone (mudstone texture close to the edges) with a massive microstructure. It is composed of 97% micro-cryptocrystalline quartz, macrocrystalline quartz, and fibrous chalcedony, found mostly replacing fossils. Other minerals include uncommon iron oxides and the porosity is of vug type (3%). Fossil content is very frequent (@fig-distance-types f2-f3), albeit mostly unidentifiable and poorly preserved. Identified fossils are rare sponge spicules, rare ostracods with some degree of preservation, common small, unidentified foraminifera (100 μm) of differing preservation degrees, and as aforementioned, foraminifera from the Pfenderenidae family (500 μm). Despite the macroscopic similarities between type 8 and cherts from the Middle Subbetic region chert formations (@fig-map), the geological reference samples showed no macroscopically visible foraminifera. However, our results indicate that despite no foraminifera being visible macroscopically, they are present in thin section. The fact that no foraminifera were identified in the small number of analysed samples from the Cadiz lithotheque may simply reflect their uncommon visibility, possibly due to patination. Another possible source may be the Upper Cretaceous wackestone/packstone cherts with foraminifers from the Lisbon region [@jordaoFlintSourcesMobility2022]. However, these studies do not specify the existence of Pfenderenidae foraminifers. Given the lack of this specific fossil in the reference samples from the Andalusian chert \~250 km east and southeast, and the central Portugal cherts 215-250 km north, the probable source of Type 8 remains unknown.

Type 9 (@fig-distance-types g) corresponds to a banded micro-cryptocrystalline quartz/micrite peloidal packstone. It corresponds to less than \~2% of the overall cherts in the Terrace and Shelter areas. It is opaque, has a heterogeneous structure and is characterized by a variable colour distribution: it shows a horizontal laminated and finely laminated pattern, with grey dark bands intercalated with light grey bands (@fig-distance-types g1). Whenever present, the cortex is rounded (FIG) and is generally thin. In specific samples, the cortex is present on two parallel planes of the sample (parallel to the laminations), which indicates the chert was originally available in bedded layers. Petrographically this chert is a banded packstone (banding of sedimentary origin; @fig-distance-types g2), composed mostly of micro-cryptocrystalline quartz (48%), micrite (40%), dolomite (10%), and accessory fibrous chalcedony and mica (muscovite). The allochems are peloids, unidentifiable bioclasts, iron oxides, and opaque minerals. Porosity is low (1%). Although its origin is unknown, when visiting the Cadiz lithotheque, we identified a small type of banded black cherts without noticeable fossils or inclusions. Despite the lack of comparative petrographic data, the macroscopic resemblances raise the possibility that Type 9 belongs to Andalusian chert formations in southern Spain, 250 km east from the site (@fig-map).

Type 10 (@fig-distance-types h) is an oolitic packstone, massive micro-cryptocrystalline micrite/quartz from a marine environment and possibly a high-energy, shallow depositional environment, due to the type, sorting, and preservation conditions of the ooids. It represents \~1% of the overall Terrace cherts and less than 0.1% of the chert in the Shelter. Its heterogeneous structure has two types of fabrics with different macroscopic characteristics: brown (opaque) and black (sub-translucent) fabric (@fig-distance-types h1). The sample chosen for the thin section is composed of 35% micro-cryptocrystalline quartz, 53% micrite, 9% macrocrystalline quartz, and accessory fibrous chalcedony (filling fractures) and mica; the percentages of quartz and micrite might be related to the fabrics present in the artefacts. Allochems are common iron oxides and very frequent ooids. Bioclasts are rare, poorly preserved, and replaced by quartz. The porosity is intraparticle and vug (10%). The ooids are distributed homogeneously across the sample (although macroscopically visible in the brown fabric) and are highly abundant and concentrated. They are poorly sorted, varying between 500 and 20 μm (@fig-distance-types h2). They show round-to-elliptical shapes, and their preservation is variable: some ooids show a poorly preserved micritic structure, while others show concentric laminae structures around a round nucleus [@flugelMicrofaciesCarbonateRocks2010]. This chert shows no similarities to any consulted reference material. Despite the existence of oolitic limestones in the Algarve region correspondent to Middle and Upper Jurassic formations and often described as whiteish (Rocha et al., 1979), the presence of oolitic chert in the region is not recognized in geological and archaeological literature [@cardoso_sobre_2018; @rocha_noticia_1979; @verissimo_jazidas_2004] and the visit of outcrops mentioned in the geological cartography did not allow their identification.

Type 11 (@fig-small-types) is a banded/laminated mudstone, characterized by the presence of a conchoidal fracture (within the nodule) and a non-conchoidal, laminar fracture (at the edges of the nodule). This chert type is grey, opaque, and without identifiable allochems. This chert shows a limited presence at the archaeological site as previously mentioned, constricted to the Solutrean occupations, and corresponding to less than 0.1% of the Terrace cherts and \~2% of Shelter cherts.

```{r}
#| label: fig-small-types
#| fig-cap: Type 11 and TL cherts macroscopic variability.

knitr::include_graphics(here::here("analysis/figures/small_grid.png"))

```

Finally, the TL types (@fig-small-types) are characterized by several samples with differing characteristics from the previously established types, and without correspondence to the local cherts and chalcedony from the Algarve region previously described [@belmiroCreatingFramesReference2023]. These are composed mostly of blanks or retouched tools, rarely with cortex, and rarely the same TL type is found in different archaeological layers. A detailed description of these types can be found in the SOM (SOM). In both areas, the TL cherts compose less than 1% of the cherts.

Apart from the types common throughout the stratigraphy of the Terrace and Shelter, Type 11 and samples from the TL type also show no identifiable source. The exceptions are TL01, TL10-11 and TL15-16. Jasper is known in the South Portuguese Zone area [@oliveira_carta_1984-1], from south/central Portugal to Spain, above the Algarve (from Aljustrel to Beja; @fig-map), suggesting that the source of TL01 may be from south/central Portugal at a minimum distance of 100 km north and northeast from Vale Boi. Similarly, despite being different from the most frequent Cretaceous nodules, the macroscopic similarities between TL10 and TL11 and specific nodules from Central Portugal in the Lusolit lithotheque may indicate these cherts are from the Estremadura/Lisbon area (@fig-map a). This difference in appearance may be due to the already-mentioned variability of the nodule’s macroscopic characteristics. Finally, TL15 and TL16 show macroscopic similarities to radiolarite reference samples located at the Cadiz lithotheque, which may indicate the source of these cherts is from the Middle Subbetic region [@dominguezbellaMineralesRocasSociedades2010; @dominguez-bella_estudio_2006].

## Stratigraphical distribution

### Terrace area

Most chert types are consistently present across the Terrace stratigraphy (except for Type 11 which appears only in level 4 and in a small number of artefacts; n=2), although there are statistically significant differences in chert types used throughout the different levels (p-value \<2.22e-16, FIG.). In the Solutrean (levels 4, 4C, and 4D) and Proto-Solutrean (levels 4E and 4E/5) occupations, local cherts make up more than \~70% of the chert and 80% when considering the weight. Despite this homogeneous pattern, local types show statistically significant differences between the two technocomplexes. In the Solutrean levels, the yellow/brown cherts (Type 2) are present in the highest percentage, at \~30%. In contrast, the grey/white cherts (Type 4) become the most used chert with percentages above 35% in level 4E and \~20% in the top of level 5. For both technocomplexes, chert non-congruent with the known local sources represent percentages lower than 30%. This pattern seems to shift at the start of the Proto-Solutrean horizon (level 5) and the Gravettian occupations (levels 6 and 7); in these levels, the percentages of local cherts drop to 45-60%, while the percentages of specific non-local chert types increase significantly, such as Type 6 (increasing from \~10% during the Solutrean to 30%). However, when considering the sum of the weight, the percentages of Type 6 diminish, instead representing around 20% of the chert use, and bringing in level 6 the use of local chert to 50% in comparison to chert types of non-local sources.

```{r}
#| label: fig-type-per-level
#| fig-cap: Terrace area chalcedony and chert type frequency (a) and weight sum (b) per level, without unidentifiable samples. Only percentages superior to the 5% threshold are shown in the figure.
#| fig-width: 10

count_table <- phase_dataset %>%
  filter(RMTYPE != "INDET") %>% 
  mutate(RMTYPE = factor(RMTYPE, levels = c("TYPE 1","TYPE 2", "TYPE 3", "TYPE 4",
                                        "TYPE 5", "TYPE 6","TYPE 7", 
                                        "TYPE 8", "TYPE 9", "TYPE 10", "TYPE 11", "TL"))) %>%
  dplyr::group_by(Level, RMTYPE) %>% 
  dplyr::summarise(n = n()) %>%
  mutate(percent = n/sum(n)) %>% 
  mutate(Technocomplex = case_when(
    Level == 4 | Level == "4C" | Level == "4D" ~ "Solutrean",
    Level == "4E" | Level == "4E/5" | Level == 5 ~ "Proto-Solutrean",
    Level == 6 | Level == 7 ~ "Gravettian")) %>% 
  mutate(Technocomplex = factor(Technocomplex, levels = c("Solutrean", "Proto-Solutrean", "Gravettian")))

count_plot <- ggplot(count_table, aes(x=Level,y=percent, fill = RMTYPE)) +
  geom_bar(stat = "identity", width = 0.5) +
  facet_grid(.~Technocomplex, scales = "free_x") +
  scale_y_continuous(labels = scales::percent_format()) +
  #theme_minimal() + #type of theme
  labs(y = "Count %") +# change labels
  guides(fill=guide_legend(title=NULL)) + # remove legend title
  geom_text(aes(label=ifelse(percent >= 0.05, paste0(sprintf("%.1f", percent*100),"%"),"")),
                position=position_stack(vjust=0.5), colour="white", size=3) +
  theme(axis.title.x = element_blank()) +
  scale_fill_manual(values = c("#C74B50", "#D49B54", "#911F27", "#630A10",
                              "#FACE7F", "#845460", "#EAD3CB", "#BDC7C9",
                              "#2B4F60", "#BDD2B6", "#798777", "#FCF0C8"))

weight_table <- phase_dataset %>%
  filter(RMTYPE != "INDET" & WEIGHT != "0") %>% 
  mutate(RMTYPE = factor(RMTYPE, levels = c("TYPE 1","TYPE 2", "TYPE 3", "TYPE 4",
                                        "TYPE 5", "TYPE 6","TYPE 7", 
                                        "TYPE 8", "TYPE 9", "TYPE 10", "TYPE 11", "TL"))) %>%
  dplyr::group_by(Level, RMTYPE) %>% 
  dplyr::summarise(wsum = sum(WEIGHT)) %>% 
  mutate(percent = wsum/sum(wsum)) %>% 
  mutate(Technocomplex = case_when(
    Level == 4 | Level == "4C" | Level == "4D" ~ "Solutrean",
    Level == "4E" | Level == "4E/5" | Level == 5 ~ "Proto-Solutrean",
    Level == 6 | Level == 7 ~ "Gravettian")) %>% 
  mutate(Technocomplex = factor(Technocomplex, levels = c("Solutrean", "Proto-Solutrean", "Gravettian")))

  weight_plot <- ggplot(weight_table, aes(x = Level, y = percent, fill = RMTYPE)) +
  geom_bar(stat = "identity", width = 0.5) +
  facet_grid(.~Technocomplex, scales = "free_x") +
  scale_y_continuous(labels = scales::percent_format()) +
  #theme_minimal() + #type of theme
  labs(x = "Level", y = "Weight %") +# change labels
  guides(fill=guide_legend(title=NULL)) + # remove legend title
  geom_text(aes(label=ifelse(percent >= 0.05, paste0(sprintf("%.1f", percent*100),"%"),"")),
                position=position_stack(vjust=0.5), colour="white", size=3) +
  theme(axis.title.x = element_blank()) + 
  scale_fill_manual(values = c("#C74B50", "#D49B54", "#911F27", "#630A10",
                              "#FACE7F", "#845460", "#EAD3CB", "#BDC7C9",
                              "#2B4F60", "#BDD2B6", "#798777", "#FCF0C8"))

ggarrange(count_plot, weight_plot, labels = c("a", "b"), nrow = 2, common.legend = TRUE, legend = "bottom")

```

```{r}
#| label: chisq
#| include: false

library(corrplot)

chisq <- phase_dataset %>% 
  filter(RMTYPE != "INDET") %>% 
  filter(RMTYPE != "TYPE 11")

chisq_tab <- table(chisq$RMTYPE, chisq$Level)
chisq_matrix <- as.table(as.matrix(chisq_tab))
chisq_rm <- chisq.test(chisq_tab)
round(chisq_rm$residuals, 3)

corrplot(chisq_rm$residuals, method = "color", is.corr = FALSE)

mosaicplot(chisq_tab, shade = TRUE, las = 2, main = "RMTYPE")

library(vcd)
vcd_labels <- list(RMTYPE = c("TL", "1", "10", "2", "3", "4",
                                   "5", "6", "7", "8", "9"))
mosaic(~ Level + RMTYPE,
      gp = shading_max,
      data = chisq, 
      set_labels = vcd_labels,
      rot_labels = c(0, 0))

```

```{r}
#| label: rm-tables
#| include:  false

# Tables f Terrace and Shelter with n(%) of chert types both overall by area and by level. Perhaps going to SOM (?)
terrace_table <- phase_dataset %>%
  filter(RMTYPE != "INDET") %>% 
  select(RMTYPE, Level) %>% 
  mutate(RMTYPE = factor(RMTYPE, levels = c("TYPE 1", "TYPE 2", "TYPE 3", "TYPE 4", "TYPE 5", "TYPE 6",
                                            "TYPE 7", "TYPE 8", "TYPE 9", "TYPE 10", "TYPE 11","TL"))) %>%
  tbl_summary(include = c(RMTYPE), by = Level) %>% 
  add_overall()

S_dataset["Layer"][is.na(S_dataset["Layer"])] <- "No info"
shelter_table <- S_dataset %>% 
  filter(RMTYPE != "INDET") %>%
  filter(Layer != "D") %>% 
  select(RMTYPE, Layer) %>% 
  mutate(RMTYPE = factor(RMTYPE, levels = c("TYPE 1", "TYPE 2", "TYPE 3", "TYPE 4", "TYPE 5", "TYPE 6",
                                            "TYPE 7", "TYPE 8", "TYPE 9", "TYPE 10", "TYPE 11","TL"))) %>%
  tbl_summary(include = c(RMTYPE), by = Layer) %>% 
  add_overall()

```

```{r}
#| label: extra-graph
#| include: false

# Violin plot for RMtype per depth
phase_dataset %>%
  filter(Z != "0") %>% 
  filter(!Z >= "25") %>% 
  mutate(RMTYPE = factor(RMTYPE, levels = c("TYPE 1","TYPE 2", "TYPE 3", "TYPE 4",
                                        "TYPE 5", "TYPE 6","TYPE 7", 
                                        "TYPE 8", "TYPE 9", "TYPE 10", "TYPE 11", "TL"))) %>%
  filter(RMTYPE != "INDET") %>% 
  ggplot(aes(RMTYPE, Z, fill = RMTYPE)) +
  geom_violin(trim = FALSE, scale="width", adjust = .5) + 
  geom_point(position = position_jitter(seed = 1, width = 0.2), alpha = 0.1, color = "darkorange") +
  geom_hline(yintercept = c(24.5,24.3,24.1,23.8,23.45), colour="#BB0000", linetype="dashed") +
  labs(title="Groups per depth", 
       subtitle="Chert groups in the Terrace area by depth",
       caption="Source: mpg",
       x="",
       y="Depth") +
  stat_summary(fun.data=mean_sdl, geom="errorbar", color="black", width=0.2) +
  scale_fill_manual(values = c("#C74B50", "#D49B54", "#911F27", "#630A10",
                              "#FACE7F", "#845460", "#EAD3CB", "#BDC7C9",
                              "#2B4F60", "#BDD2B6", "#798777", "#FCF0C8")) +
  theme(axis.text.x=element_blank(),
        axis.ticks.x=element_blank())

```

When plotting the weight, it is apparent that most of the samples are located under the 0.3 kg mark, with median and bottom quartiles often falling under the 0.5 kg line (@fig-weight). Local types, such as Type 2, Type 4, and Type 5, although mostly represented by artefacts with a small mass (\~0.3 kg), show the heaviest and highest number of heavy artefacts. In comparison, other types, although showing some variation in the weight of artefacts, show smaller artefacts and their means and medians are frequently closer to the bottom quartile. As seen in Figure 3 B, non-local types show larger masses from level 5 to level 7.

```{r}
#| label: fig-weight
#| fig-cap: "Terrace boxplots of chert type weight per level."
#| #| fig-width: 10

# Weight boxplots terrace
phase_dataset %>% 
  filter(WEIGHT !="0" & !RMTYPE %in% c("INDET","GB","TL")) %>%
  mutate(RMTYPE = factor(RMTYPE, levels = c("TYPE 1","TYPE 2", "TYPE 3", "TYPE 4",
                                        "TYPE 5", "TYPE 6","TYPE 7", 
                                        "TYPE 8", "TYPE 9", "TYPE 10", "TYPE 11", "TL"))) %>%
  mutate(Technocomplex = factor(Technocomplex, levels = c("Solutrean", "Proto-Solutrean", "Gravettian"))) %>% 
  ggplot(aes(RMTYPE, WEIGHT, color = RMTYPE)) +
  geom_boxplot() +
  facet_wrap(~Technocomplex + Level) +
  stat_summary(fun.y=mean, geom="point", shape=23) +
  labs(y="Weight") +
  coord_cartesian(ylim =  c(0, 40)) +
  scale_color_manual(values = c("#C74B50", "#D49B54", "#911F27", "#630A10",
                              "#FACE7F", "#845460", "#EAD3CB", "#BDC7C9",
                              "#2B4F60", "#BDD2B6", "#798777", "#FCF0C8")) +
  theme(axis.text.x=element_blank(),
        axis.ticks.x=element_blank(), axis.title.x = element_blank())

```

### Shelter area

The same raw materials found in the Terrace area are also found in the Solutrean occupation of the Shelter, albeit with some differences (@fig-analysis_shelter). The local cherts make up between \~70-80% of the sample and are mainly composed of the brown/yellow cherts (Type 2) that frequently characterize the Lower Jurassic outcrops of the Algarve (@fig-analysis_shelter a-b). The presence of cherts not congruent with the local sources is, like level 4 in the Terrace, close to 25%, although dominated mostly by Type 6 (\~15%). Similarly to the Solutrean occupations of the Terrace, although the artefacts are generally small (lower weights), the local types show a higher variability in mass, with several outliers and heavy samples when compared to the non-local types (@fig-analysis_shelter c).

```{r}
#| label: fig-analysis_shelter
#| fig-cap: "a-b) Shelter area chalcedony and chert type frequency (a) and weight sum (b) per level, without unidentifiable samples. Only percentages superior to the 5% threshold are shown in the figure. c): Shelter boxplots of chert type weight per level."
#| #| fig-width: 10

# Type per level plots fr Shelter
## Preparing db for the barplots
s_data_plots <- S_dataset %>% # sorting missing values and filtering non-solutrean layers
  filter(RMTYPE != "INDET" & WEIGHT != "0") %>% 
  mutate_at(vars(Layer), ~replace_na(., "ND")) %>% 
  filter(CODE != "B" & Layer != "D") %>% #removing buckets (B)
  mutate(RMTYPE = factor(RMTYPE, levels = c("TYPE 1","TYPE 2", "TYPE 3", "TYPE 4",
                                        "TYPE 5", "TYPE 6","TYPE 7", 
                                        "TYPE 8", "TYPE 9", "TYPE 10", "TYPE 11", "TL")))

## Count barplot
shelter_count <- s_data_plots %>%
  dplyr::group_by(Layer, RMTYPE) %>% 
  dplyr::summarise(n = n()) %>%
  mutate(percent = n/sum(n)) %>% 
  ggplot(aes(x = Layer, y = percent, fill = RMTYPE)) +
  geom_bar(stat = "identity", width = 0.5) +
  scale_y_continuous(labels = scales::percent_format()) +
  #theme_minimal() + #type of theme
  labs(y = "Count %") +# change labels
  guides(fill=guide_legend(title=NULL)) + # remove legend title
  geom_text(aes(label=ifelse(percent >= 0.05, paste0(sprintf("%.1f", percent*100),"%"),"")),
                position=position_stack(vjust=0.5), colour="white", size=3) +
  theme(axis.title.x = element_blank()) + 
  scale_fill_manual(values = c("#C74B50", "#D49B54", "#911F27", "#630A10",
                              "#FACE7F", "#845460", "#EAD3CB", "#BDC7C9",
                              "#2B4F60", "#BDD2B6", "#798777", "#FCF0C8"))
## Weight barplot
shelter_weight <- s_data_plots %>%
  dplyr::group_by(Layer, RMTYPE) %>% 
  dplyr::summarise(wsum = sum(WEIGHT)) %>% 
  mutate(percent = wsum/sum(wsum)) %>% 
  ggplot(aes(x = Layer, y = percent, fill = RMTYPE)) +
  geom_bar(stat = "identity", width = 0.5) +
  scale_y_continuous(labels = scales::percent_format()) +
  #theme_minimal() + #type of theme
  labs(y = "Weight %") +# change labels
  guides(fill=guide_legend(title=NULL)) + # remove legend title
  geom_text(aes(label=ifelse(percent >= 0.05, paste0(sprintf("%.1f", percent*100),"%"),"")),
                position=position_stack(vjust=0.5), colour="white", size=3) +
  theme(axis.title.x = element_blank()) + 
  scale_fill_manual(values = c("#C74B50", "#D49B54", "#911F27", "#630A10",
                              "#FACE7F", "#845460", "#EAD3CB", "#BDC7C9",
                              "#2B4F60", "#BDD2B6", "#798777", "#FCF0C8"))
## Weight boxplot
shelter_box <- s_data_plots %>% 
  filter(WEIGHT !="0" & !RMTYPE %in% c("INDET","TL") & Layer != "ND") %>%
  ggplot(aes(RMTYPE, WEIGHT, color = RMTYPE)) +
  geom_boxplot() +
  facet_wrap(.~Layer, ncol = 1) +
  stat_summary(fun.y=mean, geom="point", shape=23) +
  labs(y = "Weight") +
  #geom_jitter(shape=16, position=position_jitter(0.2)) +
  coord_cartesian(ylim =  c(0, 40)) +
  scale_color_manual(values = c("#C74B50", "#D49B54", "#911F27", "#630A10",
                              "#FACE7F", "#845460", "#EAD3CB", "#BDC7C9",
                              "#2B4F60", "#BDD2B6", "#798777", "#FCF0C8")) +
  theme(axis.text.x=element_blank(),
        axis.ticks.x=element_blank(), legend.position = "none", axis.title.x = element_blank())

## Arrange plots in a single figure
gg1 <- ggarrange(shelter_count, shelter_weight, labels = c("a", "b"), nrow = 2, common.legend = TRUE, legend="bottom")
ggarrange(gg1, shelter_box, labels = c(" ", "c", ncol = 2, common.legend = TRUE, legend = "bottom"))

```

# Discussion

The macroscopic and petrographic analysis of chert and chalcedony lithic materials from the UP levels of the Terrace and Shelter areas at Vale Boi yield essential insights into the procurement of raw materials and the mobility patterns of hunter-gatherers in southwestern Iberian Peninsula. These materials serve as valuable indicators for identifying changes in mobility, as they reveal patterns of both local and non-local procurement. The dominance of local raw materials in all assemblages indicates a heavy reliance on nearby resources. However, the observed variation in the frequencies of local types may reflect differences in mobility strategies or preferences for certain nodules. The detection of non-local raw materials throughout the stratigraphy points to a persistent use of long-distance chert sources. This continuity could be attributed to a tradition of shared knowledge and culture regarding the locations of raw materials and other resources within the landscape, or the preservation of social networks over time. @cascalheira_2017 acknowledge the adoption of specific behavioural strategies by Vale Boi hunter-gatherers in response to the regional ecological conditions, which were sustained throughout the UP at the site. Practices such as grease rendering or the functional specialization of raw materials are seen as deliberate actions with significant cultural implications for future generations, influencing the adaptive choices made over time. Accordingly, chert raw materials appear to be integral to these behavioural strategies, established from the site's earliest occupations and perpetuated throughout the UP.

The local chalcedony and chert indicate that the groups were sourcing a significant portion of their raw materials within approximately a 20 km radius of the site. The direct procurement of these materials is plausible, given that known chalcedony sources are situated to the west, about 6-7 km from Vale Boi, and chert sources are found to the east, within the inland Lower Jurassic outcrops, roughly 8 km away (@fig-map b). This hypothesis is supported by the potential existence of UP occupations near these outcrops, which might reflect quarrying and chert catchment activities during prehistory [@zambujo_jazidas_1998]. Evidence of short-distance procurement includes the presence of heavy nodules and large chunks of chert from local varieties with parent rock still attached, as well as a significant number of outliers (representing larger blocks and chunks) in the weight boxplots across the stratigraphy (Figure 4). These may have been discarded due to natural fractures. It's also possible that local cherts were procured from farther distances, perhaps as part of embedded activities. The other chert sources from the Lower and Upper Jurassic periods in the western Algarve, located along the beach and cliff areas as previously mentioned, are between 15-18 km southwest of Vale Boi (@fig-map b). While our current data do not pinpoint which specific outcrops were exploited based on macroscopic and petrographic analysis, the discovery of marine resources such as shells at the site indicates that these groups visited the coast [@belmiro_gravettian-solutrean_2021; @mannePryingNewMeaning2011]. This suggests that raw materials could have been conveniently acquired in coastal vicinities without significant detours [@binfordOrganizationFormationProcesses1979; @surovellBehavioralEcologyLithic2009]

However, the discussion around local cherts becomes more complex due to the presence of similar outcrops in the eastern Algarve. Previous research has indicated that macroscopic criteria alone are insufficient to determine whether the cherts used originated from Lower Jurassic sources in the western Algarve (within approximately a 20 km radius of the site) or from Middle Jurassic sources approximately 80 km to the east (@fig-map c). This is because the cherts are similar both macroscopically and petrographically [@belmiroCreatingFramesReference2023]. A similar issue is noted by Surovell [-@surovellBehavioralEcologyLithic2009], who described a scenario at the late Pleistocene archaeological site of Krmpotich (USA), where a specific type of chert found at the site could be sourced from both local and long-distance outcrops. Surovell suggested that some of these cherts might have been imported from distant sources or previous campsites. Consequently, we cannot rule out the possibility that some of the cherts we have identified as being sourced within a 20 km radius of Vale Boi could actually be the result of long-distance procurement and regional mobility, potentially from areas approximately 80-100 km away where Middle Jurassic cherts are located. Given the similarity in the quality of the raw materials and the absence of geochemical studies, and in light of the study and model mentioned earlier, we surmise that most of these raw materials likely originate from local sources (within 20 km), although a small portion may be attributable to regional movements.

Similarly, Upper Jurassic cherts located about 90 km east of Vale Boi (@fig-map c) have not been detected in the assemblages. This absence could be attributed to the limited use and transport of these nodules from their sources to Vale Boi, potentially due to the small size of the nodules (often between 3-5 cm), the difficulty of detaching them from the parent rock, or challenges related to their accessibility and visibility, given that the known outcrops are situated inland, approximately 10 km from the present coastline.

While the regional Upper Jurassic cherts situated less than 100 km east of the site were not identified within the assemblages, the allochthonous cherts with potentially identified sources seem to originate from greater distances and in two main directions: 1) from the north, where Cretaceous cherts from Central Portugal are found (@fig-map a), along with a small amount of jasper (TL01); and 2) from the east, where peloidal/oolitic chert from western Andalusia is present (@fig-map), and possibly jasper, transported by the Guadiana river at the current border between Portugal and Spain. The scarcity of pebble cortex and the predominance of rounded cortex suggest that these cherts were likely not collected from beaches or riverbeds far from the outcrops but were possibly gathered closer to the source areas. These identified axes of long-distance movement to the north and east could be influenced by the topography, affecting mobility patterns. Indeed, formal models that seek to explain hunter-gatherer mobility have highlighted the significance of factors such as topography on the distances travelled by groups or individuals [@brantinghamMeasuringForagerMobility2006]. For instance, Quíros Castillo [-@quiroscastilloArchaeologyHistoryPeasantries2020] discusses how Copper Age groups mitigated subsistence risks by establishing exchange networks and engaging in regional mobility, as evidenced by the trade of allochthonous raw materials, despite the high energy costs and risks associated with these exchanges. Nonetheless, the author notes certain characteristics of the Iberian Peninsula that facilitate mobility, such as the low cost of traversing the plateaus and the interconnectivity of the Guadalquivir River valley with the Guadiana and Tagus rivers. Thus, the observed long-distance mobility patterns in the north/east directions by hunter-gatherers at Vale Boi might reflect the exploitation of these natural corridors for movement, explaining the substantial use of allochthonous chert. The presence of a smaller subset of allochthonous cherts of unknown origin (TL) could represent different forms of long-distance trade of finished goods and blanks, or interactions with various social networks, potentially even indicating individual networks and mobility strategies [sensu @gamblePalaeolithicSocietiesEurope1999].

The significant presence of allochthonous cherts during the Gravettian occupation at Vale Boi suggests extensive communication and mobility among the groups residing there, indicating an expansive social network. This finding diverges from the anticipated patterns. In the context of the Solutrean occupations, one would expect a high frequency of non-local raw materials due to the broadening of social networks. In contrast, the Gravettian period, especially in light of the technological findings from southern Portugal, would typically be associated with a reduced reliance on non-local raw materials.

Technological analyses of lithic assemblages have unveiled notable differences between Gravettian assemblages from central and southern Portugal, pointing to the existence of two distinct regional Gravettian traditions in southern Iberia: the Vicentine Gravettian at Vale Boi, and the Mediterranean Gravettian identified in UP sites in southern Spain [@marreirosLithicTechnologyGravettian2015; @marreirosLithicTechnologyVariability2013]. These traditions reflect adaptations to the unique local and regional ecological settings [@bradtmoller_lithic_2016] and evolving environmental conditions [@marreirosLithicTechnologyVariability2013], potentially fostering distinct ethnographic divisions among hunter-gatherer communities [@marreirosLithicTechnologyGravettian2015].

This analysis suggests that lithic technological traits may delineate different coexistence boundaries, confined to specific regional bands and aligned with the ecological dynamics of southwestern Iberia. Consequently, one might expect territorial or cultural demarcations to influence material usage patterns, with a shift towards a greater reliance on local resources and a diminished use of non-local raw materials during the Gravettian. However, the observed prevalence of allochthonous materials could reflect not merely long-distance movements within a single cultural framework but rather the result of interactions and exchanges between different groups. Such exchanges, focused on securing raw materials not locally available, underscore the complexity of prehistoric social networks.

This perspective aligns with Brantingham's [-@brantinghamMeasuringForagerMobility2006] discussion on social exchange, which posits that exchanges likely occurred between groups occupying distinct territories, who would meet regularly to trade raw materials. This model helps explain the presence of materials from distant sources at archaeological sites, illustrating the nuanced and interconnected nature of prehistoric social landscapes.

An alternative interpretation of the varying frequencies of non-local cherts during different periods at Vale Boi can be informed by the "Mean per Capita Occupation Span" model proposed by Surovell [-@surovellBehavioralEcologyLithic2009] and the "Distance/Frequency Index (DFI)" outlined by Grove et al. [@groveMovingFarMoving2023]. These models hypothesize that shorter-term occupations are likely to show a higher proportion of non-local artefacts, whereas with longer-term settlements, the prevalence of local artefacts increases. This pattern may be attributed to environmental influences: challenging conditions such as fragmented habitats and cooler temperatures limit mobility, making long-term settlement in areas rich in local resources a preferable, lower-risk option [@surovellBehavioralEcologyLithic2009]. This pattern is mirrored in the archaeological record of Vale Boi, where, aside from the earlier Gravettian and Magdalenian phases, subsequent occupations seem to adopt a logistical-settlement strategy, positioning Vale Boi as a long-term basecamp [@cascalheira_2017].

In this context, the Gravettian phase at Vale Boi could reflect a period of enhanced mobility and brief occupations, possibly interspersed with reoccupations, indicative of a significant use of non-local raw materials. This scenario parallels observations from the Late Gravettian in the French Massif Central, where hunter-gatherer groups are known to have transported entire or pre-formed blocks of raw material over considerable distances for short-term, hunting-related site occupations [@delvigneGeoresourcesTechnoculturalExpressions2019]. However, the shift toward colder and drier climatic conditions marked by Heinrich Event 2 during the Proto-Solutrean, followed by the Last Glacial Maximum in the Solutrean period [@cascalheiraHunterGathererEcodynamics2013; @gonzalez-samperiz_steppes_2010-1; @sanchez_goni_millennial-scale_2010; @schmidt_rapid_2012; @belmiro_gravettian-solutrean_2021], resulted in a more challenging environment with sparser vegetation [@cascalheiraHunterGathererEcodynamics2013; @gonzalez-samperiz_steppes_2010-1]. These changes likely constrained the mobility of hunter-gatherers, fostering longer-term settlements within the resource-rich niche of southwestern Portugal [@cascalheira_2017; @schmidt_rapid_2012]. Consequently, as settlements extended in duration, the reliance on non-local raw materials, whether sourced directly or acquired through trade, diminished in favour of increasingly dominant local raw materials.

# Conclusions

Our findings build upon earlier studies indicating that the majority of chert utilized at Vale Boi originates from local Lower Jurassic sources within a 20 km radius of the site, with non-local materials constituting a minor portion of the chert used [@pereira_territory_2016-1]. The hypothesis of long-distance procurement of Cretaceous cherts from central Portugal, situated north of Vale Boi, is corroborated by visual and petrographic evidence [@pereira_territory_2016-1; @bicho_ecodynamics_2013], and this pattern persists throughout the UP. Furthermore, the identification of cherts bearing macroscopic resemblance to the peloidal/oolitic Upper Jurassic cherts from the Betic System in southern Spain confirms a west-east mobility axis, previously inferred from techno-typological studies, spanning the entirety of the UP period.

These findings reveal a landscape of hunter-gatherers engaged in complex practices of landscape exploration, resource management, and possibly risk management. The presence of allochthonous raw materials, albeit in small quantities during the Proto-Solutrean and Solutrean phases, underscores that the inhabitants of Vale Boi—situated on the periphery of the Iberian Peninsula and Europe, bordered by the sea to the south and west—were not isolated. Instead, they were part of an extensive social network, a pattern of interconnectivity that resonates with findings across western Europe. In southern Portugal, this interconnection among UP hunter-gatherers from diverse and distant regions of the Iberian Peninsula is similarly observed.

The outcomes of this research pave the way for continued exploration of these assemblages through diverse methodologies. Future directions include employing a geochemical approach to both the studied assemblages and the geological reference collection, aiming to enrich discussions on mobility throughout the UP at Vale Boi. This approach is anticipated to provide diagnostic elements to distinguish between cherts from Lower and Middle Jurassic units, thereby offering a comprehensive view of regional resource utilization and mobility patterns.

Additionally, experimental replication and the creation of a reference collection modified through heat treatment could enhance the identification of raw material types and currently indistinguishable samples, as noted in prior research [@pereira_territory_2016-1]. Lastly, we intend to integrate technological data from current and ongoing datasets to delve deeper into the usage and management of raw materials (both local and non-local) over time, their association with different site uses, and to contextualize these practices within the broader framework of human adaptations to changing environmental conditions throughout the UP.

# Acknowledgements

We would like to thank the Fundacão para a Ciência e Tecnologia for the funding essential for our research, and ICArEHB for funding and providing access to the archaeological materials. We would also like to thank Dr. David Ortega for the provided insight and expertise that greatly assisted the research and Dr. Célia Gonçalves for the invaluable assistance with the maps and imagery created for this paper.

# Declarations

## Funding

This study was supported by the Fundacão para a Ciência e Tecnologia (https://www.fct.pt/), attributed to Joana Belmiro with the contract reference 2020.08722.BD (DOI: 10.54499/2020.08722.BD), and João Cascalheira with the contract reference DL57/2016/CP1361/CT0026. This study was also supported by the Seed Fund’21 attributed to Joana Belmiro, by ICArEHB – The interdisciplinary Center for Archaeology and Evolution of Human Behaviour, funded by the Fundacão para a Ciência e Tecnologia (https://www.fct.pt/) under program UIDP/04211/2020. Open Access was funded by Fundacão para a Ciência e Tecnologia (https://www.fct.pt/) and Universidade do Algarve.

## Competing Interests

The authors have no competing interests to declare that are relevant to the content of this article.

## Data availability

The authors confirm that all data generated and analysed during this study are available to the public in the Open Science Framework repository at: ++.

## Author Contributions

Study conception and design was created by Joana Belmiro, João Cascalheira and Xavier Terradas. Material preparation, data collection and analysis were performed by Joana Belmiro. The first draft of the manuscript was written by Joana Belmiro and all authors commented on previous versions of the manuscript. All authors read and approved the final manuscript.

# Supplementary Information

**Online Resource 1**: Data dictionaries for macroscopic (group and individual) and petrographic analyses.

**Online Resource 2**: Terrace and shelter lithotypes complete macroscopic and petrographic description.

**Online resource 3**: Summarised type macroscopic and petrographic features and epoch/stage identification.

<!-- The following line inserts a page break  -->

\newpage

# References

<!-- The following line ensures the references appear here for the MS Word or HTML output files, rather than right at the end of the document (this will not work for PDF files):  -->

::: {#refs}
:::

\newpage

### Colophon

This report was generated on `r Sys.time()` using the following computational environment and dependencies:

```{r}
#| label: colophon
#| cache: false

# which R packages and versions?
if ("devtools" %in% installed.packages()) devtools::session_info()
```

The current Git commit details are:

```{r}
# what commit is this file at? 
#if ("git2r" %in% installed.packages() & git2r::in_repository(path = ".")) git2r::repository(here::here())  
```