### 2024-06-28 analysis/supplementary-materials/SOM_chisquare.qmd

---
title: "Chert use and procurement patterns during the Upper Palaeolithic in Southwesternmost Iberia, Journal of Paleolithic Archaeology"
subtitle: "Online Resource 4: Statistical analysis of Terrace chert types"
author:
  - Joana Belmiro:
      orcid: https://orcid.org/0000-0003-4951-2907
      correspondence: "yes"
      email: jfbelmiro@ualg.pt
      institute:
        - Interdisciplinary Center for Archaeology and the Evolution of Human Behaviour, University of Algarve, Faro, Portugal
  - Xavier Terradas:
      orcid: https://orcid.org/0000-0002-8000-5607
      institute: 
        - Institución Milá y Fontanals de Investigación en Humanidades, Consejo Superior de Investigaciones Científicas, Barcelona, Spain
  - Salvador Dominguez-Bella:
      orcid: https://orcid.org/0000-0003-3892-763X
      institute: 
        - UGEA-PHAM, Departamento de Ciencias de la Tierra, Facultad de Ciencias, Universidad de Cádiz. Puerto Real, Cádiz, Spain
  - João Cascalheira:
      orcid: https://orcid.org/0000-0003-0321-8892
      institute:
        - Interdisciplinary Center for Archaeology and the Evolution of Human Behaviour, University of Algarve, Faro, Portugal
title-block-published: "Last updated"  
format: 
  docx:
    reference-doc: "../templates/template.docx" # Insert path for the DOCX file
execute:
  echo: false
  warning: false
  message: false
  comment: "#>"
  fig-path: "../figures/"
  fig-dpi: 600
number-sections: true
filters:
  - ../templates/scholarly-metadata.lua
  - ../templates/author-info-blocks.lua
  - ../templates/pagebreak.lua
bibliography: ../paper/references.bib
toc: false
csl: "../templates/journal-of-archaeological-science.csl" # Insert path for the bib-style
---

```{r}
#| label: callpackages

# Call packages
# This includes all necessary packages to run the code
library(readr)
library(dplyr)
library(stringr)
library(tidyr)
library(knitr)
library(corrplot)
library(vcd)
```

```{r}
#| label: db-setup

# Terrace DB preparation (repeated from main document)
# Read DBs
T_RM_data <- read_csv(here::here('analysis/data/raw_data/MACRO_IND_TERRACE.csv')) # Terrace RM analysis
context <- read_csv(here::here('analysis/data/raw_data/VB_Terrace_context.csv')) # Terrace Field DB - lithic artefact info
xyz <- read_csv(here::here('analysis/data/raw_data/VB_Terrace_xyz.csv')) # Terrace Field DB - just coordinates

# Df cleaning and uniformizing
cols_to_concat <- c("UNIT", "ID") # Vector to concatenate Unit and ID in one single variable (e.g. A6-100) - to use with both Field databases
context$ID <- str_squish(context$ID) # Remove empty spaces from ID field
xyz$ID <- str_squish(xyz$ID) # Remove empty spaces from UNIT field

# Concatenate UNIT and ID variables in one single variable (e.g. A6-100)
context <- context %>%
  dplyr::rename("UNIT" = "Unit") %>% # Renaming first column
  unite(col='ID', cols_to_concat, sep="-", remove=FALSE) %>% # Unite Unit with Id s«using cols_to_concat
  select(ID, Level, Spit, Code) %>% # Selecting important variables
  distinct(ID, .keep_all = TRUE)

xyz <- xyz %>%
  unite(col='ID', cols_to_concat, sep="-", remove=FALSE) %>% # Unite Unit with Id s«using cols_to_concat
  filter(Suffix == 0) %>% # To avoid several entries related to several points of the same artefact
  select(ID, X, Y, Z) %>% # Selecting important variables
  distinct(ID, .keep_all = TRUE)

field_data <- full_join(xyz, context, by = "ID") # Join field tables
T_dataset <- left_join(T_RM_data, field_data, by = "ID") # Join field and RM type tables

T_dataset <- T_dataset %>% 
  filter(!Level %in% c("2", "3")) %>%
  drop_na(Level) %>% 
  mutate(Level = case_when(Level == "7C" ~ "7",
                           Level == 8 ~ "7", # Collapsing samples from Level 7C and Level 8 into 7, since 8 disappears in field_data
                           Level == "4B" ~ "4", # collapsing samples from 4B to 4 due to small number
                           TRUE ~ Level)) %>% 
  mutate(RMTYPE = case_when(RMTYPE == "Chalcedony" ~ "TYPE 1", # Altering first group names to final group names. Subtypes are maintained as before.
                            RMTYPE == "TYPE 15" ~ "TYPE 3",
                            RMTYPE == "TYPE 7" ~ "TYPE 4",
                            RMTYPE == "TYPE 12" ~ "TYPE 5",
                            RMTYPE == "TYPE 11" ~ "TYPE 6",
                            RMTYPE == "TYPE 13" ~ "TYPE 7",
                            RMTYPE == "TYPE 11F" ~ "TYPE 8",
                            RMTYPE == "TYPE 13B" ~ "TYPE 9",
                            RMTYPE == "Oolitic" ~ "TYPE 10",
                            RMTYPE == "GB" ~ "TYPE 11",
                            TRUE ~ RMTYPE))

# Mutate level 5 to add phases: top of 5 (4E/5) and bottom of 5 (5), following Belmiro et al. (2020).
phase_dataset <- T_dataset %>% 
  mutate(Level = case_when(
    Level == 5 & Z > 24.1 ~ "4E/5",
    Level == 5 & Z < 24.1 ~ "5",
    TRUE ~ Level)) %>% 
  mutate(Technocomplex = case_when(
    Level == 4 | Level == "4C" | Level == "4D" ~ "Solutrean",
    Level == "4E" | Level == "4E/5" | Level == 5 ~ "Proto-Solutrean",
    Level == 6 | Level == 7 ~ "Gravettian"))
```

```{r}
#| label: chisq
#| include: false

chisq <- phase_dataset %>% 
  filter(RMTYPE != "INDET") %>% 
  filter(RMTYPE != "TYPE 11")

chisq_tab <- table(chisq$RMTYPE, chisq$Level)
chisq_matrix <- as.table(as.matrix(chisq_tab))
chisq_rm <- chisq.test(chisq_tab)
round(chisq_rm$residuals, 3)

```

The chert raw material analysis of the Terrace and Shelter areas of the archaeological site of Vale Boi revealed the presence of varied types of local and non-local chert and chalcedony throughout the stratigraphy. However, we identified differences in the percentages of these types throughout the several Upper Paleolithic occupations in the Terrace area.

To explore the statistical significance of these patterns in the Terrace area, a chi-square test of independence was performed to determine whether the chert types are related to the archaeological layer. The results are significant, X-square = 817.79, *p* \< 2.2e-16, showing statistically significant differences in the chert types present throughout the occupations. @fig-mosaicplot shows a mosaic plot visualising the residuals of the test and @fig-corrplots shows the visual correlations between the variables in level and chert type. These show different associations and patterns throughout the different levels and occupations.

In the Solutrean occupations, Type 2 and Type 3 cherts are relatively more frequent in Level 4, and although to a lesser degree, Type 5 is relatively more frequent in Level 4C. For the Solutrean occupations, all positive correlations relate to local cherts.

For the Proto-Solutrean occupations, Type 4 cherts are relatively more frequent in Levels 4E and 4E/5, and Type 1 chalcedony is relatively more frequent in Level 4E/5. Similar to the Solutrean levels, only local cherts and chalcedony show positive correlations. In comparison, Type 6 and 7 are relatively more frequent in Level 5, while Type 4 is relatively less frequent than expected. This shows a significant shift from positive correlations of local cherts to non-local.

The Gravettian occupations, especially Level 6, show markedly different patterns. TL cherts, Type 10, Type 6 and Type 9 cherts (all non-local) are relatively more frequent in Level 6. The same patterns can also be seen in @fig-corrplots, although there are expanded (although weaker) correlations between non-local chert types and Level 6 and 7.

```{r}
#| label: fig-mosaicplot
#| fig-cap: Mosaic plot showing standardized Peason's residuals from the chi-square test showing the dependencies of the different archaeological levels and chert types.

vcd_labels <- list(RMTYPE = c("TL", "1", "10", "2", "3", "4",
                                   "5", "6", "7", "8", "9"))
mosaic(~ Level + RMTYPE,
      gp = shading_max,
      data = chisq, 
      set_labels = vcd_labels,
      rot_labels = c(0, 0))
```

```{r}
#| label: fig-corrplots
#| fig-cap: Correlation plot using Pearson's residuals from the chi-square test, showing positive and negative correlations between archaeological levels and chert types.

corrplot(chisq_rm$residuals, method = "color", is.corr = FALSE)

```