### 2024-10-23 analysis/supplementary-materials/SOM_datadictionaries.qmd
Added variables to better correspond to the other SI documents

---
title: "Within and beyond: chert procurement patterns during the Upper Palaeolithic in Southwesternmost Iberia"
subtitle: "Online Resource 1: Data dictionaries for macroscopic and petrographic analyses"
author:
  - Joana Belmiro:
      orcid: https://orcid.org/0000-0003-4951-2907
      correspondence: "yes"
      email: jfbelmiro@ualg.pt
      institute:
        - Interdisciplinary Center for Archaeology and the Evolution of Human Behaviour, University of Algarve, Faro, Portugal
  - Xavier Terradas:
      orcid: https://orcid.org/0000-0002-8000-5607
      institute: 
        - Institución Milá y Fontanals de Investigación en Humanidades, Consejo Superior de Investigaciones Científicas, Barcelona, Spain
  - Salvador Dominguez-Bella:
      orcid: https://orcid.org/0000-0003-3892-763X
      institute: 
        - UGEA-PHAM, Departamento de Ciencias de la Tierra, Facultad de Ciencias, Universidad de Cádiz. Puerto Real, Cádiz, Spain
  - João Cascalheira:
      orcid: https://orcid.org/0000-0003-0321-8892
      institute:
        - Interdisciplinary Center for Archaeology and the Evolution of Human Behaviour, University of Algarve, Faro, Portugal
title-block-published: "Last updated"  
format: 
  docx:
    classoption: landscape
    reference-doc: "../templates/template_landscape.docx" # Insert path for the DOCX file
execute:
  echo: false
  warning: false
  message: false
  comment: "#>"
  fig-path: "../figures/"
  fig-dpi: 600
number-sections: true
filters:
  - ../templates/scholarly-metadata.lua
  - ../templates/author-info-blocks.lua
  - ../templates/pagebreak.lua
bibliography: ../paper/references.bib
toc: true
csl: "../templates/journal-of-archaeological-science.csl" # Insert path for the bib-style
---

# Macroscopic type characterization

```{r}
#| label: tbl-macrogroup
#| tbl-cap: Data dictionary for macroscopic characterization of the archaeological chert types and sub-types.

tibble::tribble(
  ~Variable, ~"Allowed values", ~Description, ~Reference,
  "Type", "-", "Chert type ID.", "-",
  "Sub-type", "-", "Chert subtype ID.", "-",
  "Color", "-", "Color according to the Munsell Soil Color Chart.", "-",
  "Fabric", "Homogeneous, Heterogeneous", "Distribution of color, luster, translucency, feel and pattern in the sample. **Heterogeneous**: presence of different characteristics within one single sample.", "@luedtke_archaeologists_1992",
  "Luster", "Shiny, Medium, Dull", "The character of light reflected by minerals. **Shiny**: very intense luster; **Medium**: some luster but it is not very intense; **Dull**: no luster.", "@crandell_macroscopic_2005-1 and @luedtke_archaeologists_1992",
  "Translucency - qualitative", "Highly translucent, Translucent, Sub-translucent, Opaque", "Degree to which the light can penetrate a material when held 5cm from a 40w lamp. **Highly translucent**: transparent or near transparent; **Translucent**: silhouettes can be seen through the sample’s thin part and/or light passes through the thick parts; **Sub-translucent**: light only passes through the thin parts; **Opaque**: no or almost no light passes through.", "@bressyCaracterisationGestionSilex2002 and @crandell_macroscopic_2005-1",
  "Translucency - quantitative", "-", "Measurement in mm of the thickness where light begins to pass through in translucent to sub-translucent samples.", "-",
  "Feel", "Smooth, Semi-smooth, Rough", "A fingernail is dragged across the sample’s surface. **Smooth**: nothing can be felt; **Semi-smooth**: slight rough feel; **Rough**: if the sample is distinctly rough.", "@bressyCaracterisationGestionSilex2002 and @crandell_macroscopic_2005-1",
  "Pattern", "Shaded, Spotted, Splotched, Broad mottling, Marbled mottling, Speckling, Flecks, Banded, Streaked, Laminated, Finely laminated", "Type of pattern present on the sample. **Shaded**: gradient color or texture change; **Spotted**: circle patterns less than 30% of the surface area; **Splotched**: irregular shapes less than 30% of the surface area; **Broad mottling**: large, irregular blotching, covering more than 30% of the surface; **Marbled mottling**: large relatively round shapes, covering more than 30% of the surface; **Speckling**: small dots, well distributed over the surface; **Flecks**: small dots, grouped together; **Banded**: regular lines thicker than 1 cm; **Streaked**: less regular and wider banding; **Laminated**: lines with thickness inferior to 1 cm; **Finely laminated**: series of lines with thickness inferior to 1 mm.", "@crandell_macroscopic_2005-1",
  "Cortex type", "Outcrop, Rounded, Pebble, Indeterminate", "**Outcrop**: uneven and low level of alteration to the original cortical surface; **Rounded**: homogeneous and rounded cortex, with possible slight alteration to the original cortical surface; **Pebble**: Smooth and highly weathered surface, with high degree of alteration to the cortical surface.", "-",
  "Cortex maximum thickness", "-", "Measurement in mm of the thickest cortex area.", "-",
  "Cortex minimum thickness", "-", "Measurement in mm of the thinnest cortex area.", "-",
  "Cortex transition", "Sharp, Gradual", "**Sharp**: transition between cortex and chert is sharp, with an obvious change in colors and textures; **Gradual**: cortex and chert show a gradual transition with gradient colors and textures", "-",
  "Morphology of the nodules", "Irregular or rounded nodules, tabular, bedded, lenses", "Inferred morphology of the chert based on cortical samples, unknapped chunks or preserved geological features such as encasing rock attached.", "@trewinSEDIMENTARYROCKSChert2005",
  "Allochems", "-", "Macroscopically visible allochems (excluding bioclasts).", "-",
  "Bioclasts", "-", "Macroscopically visible bioclasts.", "-",
  "Fracture", "Conchoidal, Uneven", "Presence or absence of a conchoidal fracture.", "-",
  "Quality", "Good, Acceptable, Low", "Combination of: 1) homogeneity, understood as the lack of fractures, flaws and cracks; 2) fine granularity and smooth feel; 3) lack of inclusions that due to their nature or distribution impact the knapping.", "@luedtke_chert_1978 and @brandl_genesis_2013"
) |>
  knitr::kable()

```

# Macroscopic individual analysis

```{r}
#| label: tbl-macroind
#| tbl-cap: Data dictionary for macroscopic analysis of archaeological chert artifacts.

tibble::tribble(
  ~Variable, ~"Allowed values", ~Description, ~Reference,
  "Area", "Terrace, Shelter", "Archaeological artifact provenance area from Vale Boi.", "-",
  "Level", "-", "Individual artefact archaeological level.", "-",
  "ID", "-", "Individual artefact ID.", "-",
  "Type", "-", "Type of chert characterized with macroscopic and microscopic methods.", "-",
  "Sub-type", "-", "Subtype of chert characterized with macroscopic and microscopic methods.", "-",
  "Weight", "-", "Weight of individual artefacts measured in gr.", "-",
  "Alterations", "Cracks, Potlids, Reddening, Gloss, White patina, Patina, Other", "**Cracks**: surface shrinkage cracks or crazing, described as the result of contact to high temperatures; **Potlids**: small, circular and concave negative where a fragment appears to have popped off; **Reddening**: changes to the surface of the chert from the original color to a reddish/brown color; **Gloss**: glossy luster which differs from the original chert's luster, present in all the artifacts or in specific removals; **White patina**: film, rind or layer produced on the surface of the chert visible as a color change from the original color to a white color; **Patina**: film, rind or layer produced on the surface of the chert visible as a color change from the original color (excluding white patina); **Other**: other types of alterations to the original appearance of chert, such as rolled edges or pits.", "@hurst_patination_1961; @glauberman_flint_2012; @luedtke_chert_1978; @abdolahzadehHeatExposedLithics2023; @pattersonThermalDamageChert1995; @purdyFracturesArchaeologist1975",
  "Cortex", "Yes, No", "Presence or absence of cortex.", "-",
  "Cortex thickness", "Thin, Medium, Thick", "Thickness of cortex measured in cm, following defined intervals. **Thin**: cortex thickness equal or inferior to 0.5 cm; **Medium**: 0.5-2 cm; **Thick**: >2 cm.", "-",
  "Cortex appearance", "Outcrop, Rounded, Pebble, Indeterminate", "**Outcrop**: uneven and low level of alteration to the original cortical surface; **Rounded**: homogeneous and rounded cortex, with possible slight alteration to the original cortical surface; **Pebble**: Smooth and highly weathered surface, with high degree of alteration to the cortical surface", "-",
  "Inclusions", "-", "Macroscopically visible inclusions (allochems and bioclasts).", "-"
) |>
  knitr::kable()
```

# Petrographic analysis

```{r}
#| label: tbl-petrography
#| tbl-cap: Data dictionary for petrographic characterization of thin sections.

library(tibble)

tibble::tribble(
  ~Variable, ~"Allowed values", ~Description, ~Reference,
  "ID", "-", "Individual thinsection ID.", "-",
  "Outcrop/level", "-", "Outcrop name or archaeological level.", "-",
  "Lithology", "-", "Type of rock.", "-",
  "Texture", "Mudstone, Wackestone, Packstone, Grainstone, Boundstone, Other", "**Mudstone**: muddy carbonate rock containing less then 10 % grains; **Wackestone**: mud-supported carbonate rock containing more than 10 % grains; **Packstone**: grain-supported muddy carbonate rock; **Grainstone**: mud-free carbonate rocks, which are grain supported; **Boundstone**: carbonate rocks showing signs of being bound during deposition.", "@dunhamClassificationCarbonateRocks1962",
  "Microstructure", "Homogeneous, Banded, Laminar, Nodular, Brechoid, Other", "Distribution of crystals and clasts within the rock at a microscopic scale. **Homogeneous**: equally spread in the rock; **Banded**: distributed in bands; **Nodular**: distributed in clumps; **Brechoid**: fracturing of the rock irregularly.", "@castro_dorado_petrografibasica_1989",
  "Orthochem", "-", "Materials formed in two ways: 1) deposited directly from supersaturated aqueous solutions due to chemical reactions or evaporation; 2) formed by the replacement of existing sedimentary materials.", "@vernonMicrostructuresSedimentaryRocks2018",
  "Orthochem type", "Essential (ES), Accessory (AC), Secondary (SE)", "**Essential**: minerals that form more than 5% of the volume of the rock; **Accessory**: Minerals with proportion of less than 5% of the volume of the rock; **Secondary**: products of the alteration (hydrothermal or physical), independent of the proportion within the rock.", "@castro_dorado_petrografibasica_1989",
  "Orthochem description", "-", "General description of the orthochem and where it is identified.", "-",
  "Orthochem (%)", "-", "Approximate percentage of the orthochem’s presence in the total thin-section area.", "-",
  "Allochem", "-", "Material formed by the movement and reorganization into new shapes by chemical, physical or biological processes within the depositional basin (ex. ooliths, fecal pellets, iron oxide minerals).", "@vernonMicrostructuresSedimentaryRocks2018",
  "Allochem (freq)", "Rare, Uncommon, Common, Very frequent", "**Rare**: present one or two elements; **Uncommon**: present three to 10 elements; **Common**: present 11 to 20 elements; Very frequent:> 20 elements.", "-",
  "Bioclast", "-", "Also known as skeletal particles, are the remains (complete or fragmented) of the hard parts of carbonate-secreeting organisms.", "@adamsAtlasSedimentaryRocks1988",
  "Bioclast (freq)", "Rare, Uncommon, Common, Very frequent", "**Rare**: present one or two elements; **Uncommon**: present three to 10 elements; Common: present 11 to 20 elements; **Very frequent**:> 20 elements.", "-",
  "Porosity (%)", "-", "Approximate frequency of effective porosity.", "-",
  "Porosity type", "Interparticle, Moldic, Fenestral, Fracture, Vuggy, Shelter, Other", "**Interparticle**: porosity between particles; **Moldic**: Porosity formed by selective removal of an individual constituent of the rock; **Fenestral**: Pores larger than grain-supported interstices (interparticle); **Fracture**: Porosity formed by fracturing; **Vug**: Pores larger then 1/16 mm in diameter and somewhat equant in shape; **Shelter**: Porosity created by the sheltering effect of large sedimentary particles; **Other**: Other types of porosity.", "@choquetteGeologicNomenclatureClassification1970",
  "Sedimentary structures", "Parallel lamination, Convoluted lamination, Bands/zonations, Burrow, Other", "**Parallel lamination**: sedimentary strata less than 10 mm thick, recognizable due to variation in structure or composition and more or less parallel bounding surfaces; **Convoluted lamination**: symmetrical about a vertical plane or leaning and asymmetrical, and usually exhibit narrow vertical upturned laminae, often truncated at the top, separated by a broader synclinal downfolds; **Bands/zonations**: limited areas with different characteristics related to changes in the sedimentation or cementation process; **Burrow**: bioturbation structures caused by activity of an organism that disrupts the stratification features; **Other**: other types of sedimentary structures.", "@middletonEncyclopediaSedimentsSedimentary2005"
  ) |>
knitr::kable()
```

# References